
---
output: html_document
editor_options: 
  chunk_output_type: console
---

Project description: Finding germline variants from childhood with rhabdomyosarcoma, and identifying ancestral origins of these variants

```{bash GATK pipeline on exosome sequencing data}

# raw files are saved at/project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832

cd /project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832


```

```{bash install tools}

############################

# TrimGalore
## low-quality base calls are trimmed off from the 3' end of the reads before adapter removal
## Cutadapt finds and removes adapter sequences from the 3’ end of reads.
## filter trimmed reads based on their sequence length (default: 20 bp) since quality and/or adapter trimming may result in very short sequences

module load gcc/11.3.0 py-cutadapt fastqc/0.11.9

cd /project/wiemels_260/sebastian/dependencies/trimGalore

## Check that dependents are installed
cutadapt --version
fastqc -v

## Install Trim Galore
#curl -fsSL https://github.com/FelixKrueger/TrimGalore/archive/refs/tags/0.6.7.tar.gz -o trim_galore.tar.gz
curl -fsSL https://github.com/FelixKrueger/TrimGalore/archive/refs/tags/0.6.6.tar.gz -o trim_galore.tar.gz

tar xvzf trim_galore.tar.gz

#export PATH="/project/wiemels_260/sebastian/dependencies/trimGalore/TrimGalore-0.6.6/:$PATH"

## Run Trim Galore
module load gcc/11.3.0 py-cutadapt fastqc/0.11.9
trim_galore

############################

```

> start from here if want to try trim_galore

```{bash trim_galore.sh}

#!/bin/bash
#SBATCH --time=30:00:00
#SBATCH --mail-type=ALL
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/trimgalore_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/trimgalore_%A_%a.errcd
#SBATCH --job-name=trimgalore
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=45G
#SBATCH --partition=main

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp.sh
module load gcc/11.3.0 py-cutadapt fastqc/0.11.9

dir=(`ls -d /project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832/Sample* | sed -n ${SLURM_ARRAY_TASK_ID}p`)
fas1=(`ls $dir | grep '_R1_'`)
fas2=(`ls $dir | grep '_R2_'`)

fastq1=(`echo ${dir}/${fas1}`)
fastq2=(`echo ${dir}/${fas2}`)

trim_galore --illumina --paired --fastqc -o trim_galore/ $fastq1 $fastq2 

```

> start from here if want to try fast qc only

```{bash fastqc}

#!/bin/bash
#SBATCH --time=30:00:00
#SBATCH --mail-type=ALL
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/fastqc_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/fastqc_%A_%a.errcd
#SBATCH --job-name=fastqc
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=45G
#SBATCH --partition=main

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/fastqc
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp.sh
module load gcc/11.3.0 fastqc/0.11.9

dir=(`ls -d /project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832/Sample* | sed -n ${SLURM_ARRAY_TASK_ID}p`)
fas1=(`ls $dir | grep '_R1_'`)
fas2=(`ls $dir | grep '_R2_'`)

fastq1=(`echo ${dir}/${fas1}`)
fastq2=(`echo ${dir}/${fas2}`)

fastqc --noextract $fastq1 $fastq2 \
  --outdir /project/lishaobo_902/sebastian/rhabdo/snp_dis/fastqc\
  -t 20

```

create unaligned bam

```{bash fast2ubam.sh}

#!/bin/bash
#SBATCH --time=30:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/fast2ubam_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/fast2ubam_%A_%a.errcd
#SBATCH --job-name=fast2ubam
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=45G
#SBATCH --partition=main 

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/fast2ubam
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp2.sh

dir=(`ls -d /project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832/Sample* | sed -n ${SLURM_ARRAY_TASK_ID}p`)
fas1=(`ls $dir | grep '_R1_'`)
fas2=(`ls $dir | grep '_R2_'`)

fastq1=(`echo ${dir}/${fas1}`)
fastq2=(`echo ${dir}/${fas2}`)
base=$(basename ${fastq1} _R1_001.fastq.gz)

OUTFILE=/project/lishaobo_902/sebastian/rhabdo/snp_dis/fast2ubam/${base}_unaligned_read_pairs.bam

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk FastqToSam \
    --java-options "-Xmx38G"\
    -F1 $fastq1 \
    -F2 $fastq2 \
    -O $OUTFILE \
    -SM $base \
    -RG "${base}_CBDAFANXX" \
    -PG PGS-CORE_DNA_v1.4.2 \
    -PU CBDAFANXX \
    -PL ILLUMINA \
    -LB $base 
    
#singularity run \
#  --bind /project/lishaobo_902/sebastian/rhabdo/snp_dis:/snp_dis \
#  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif
  
```

create aligned bam

```{bash bwa}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/bwa_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/bwa_%A_%a.errcd
#SBATCH --job-name=bwa
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main

module load gcc/11.3.0 bwa/0.7.17

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/bwa
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp1.sh

ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa

dir=(`ls -d /project/lishaobo_902/sebastian/rhabdo/sample_dir_000007832/Sample* | sed -n ${SLURM_ARRAY_TASK_ID}p`)
fas1=(`ls $dir | grep '_R1_'`)
fas2=(`ls $dir | grep '_R2_'`)
fastq1=(`echo ${dir}/${fas1}`)
fastq2=(`echo ${dir}/${fas2}`)

base=$(basename ${fastq1} _R1_001.fastq.gz)

outfile=/project/lishaobo_902/sebastian/rhabdo/snp_dis/bwa/${base}_aligned.bam

bwa mem $ref \
  -R "@RG\tID:${base}_CBDAFANXX\tPL:ILLUMINA:CBDAFANXX\tLB:${base}\tSM:${base}\tPG:PGS-CORE_DNA_v1.4.2" \
  -t 20 $fastq1 $fastq2 |\
  samtools view -S -b > $outfile


```

merge unaligned and aligned bam files

```{bash}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/mergebam_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/mergebam_%A_%a.errcd
#SBATCH --job-name=mergebam
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/mergebam
ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa

BAM=(`ls /project/lishaobo_902/sebastian/rhabdo/snp_dis/bwa/*aligned.bam | sed -n ${SLURM_ARRAY_TASK_ID}p`)

base=$(basename ${BAM} _aligned.bam)

UBAM=/project/lishaobo_902/sebastian/rhabdo/snp_dis/fast2ubam/${base}_unaligned_read_pairs.bam

O=${base}_merged_alignments.bam

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk MergeBamAlignment \
    --java-options "-Xmx38G"\
    -ALIGNED $BAM\
    -UNMAP_CONTAMINANT_READS true\
    -CLIP_ADAPTERS false\
    -MAX_RECORDS_IN_RAM 5000000\
    -UNMAPPED $UBAM\
    -O $O \
    -R $ref \
    -CREATE_INDEX true 

```

from mark duplicates to HaplotypeCaller
keren's refs are saved at: /project/desmith_488/kerenxu/refs

```{bash Caller}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/caller_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/caller_%A_%a.errcd
#SBATCH --job-name=caller
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main


cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller

ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa
known_sites=/project/wiemels_260/sebastian/dependencies/gatk/refs/All_20180418-3.vcf.gz
BED=/project/wiemels_260/sebastian/dependencies/gatk/refs/ace_liftover_r5.bed

FILE=(`ls /project/lishaobo_902/sebastian/rhabdo/snp_dis/mergebam/*_merged_alignments.bam | sed -n ${SLURM_ARRAY_TASK_ID}p`)
echo ${FILE}

base=$(basename ${FILE} _merged_alignments.bam)

OUTFILE=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}.marked_duplicates.bam
MARKFILE=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}.marked_dup_metrics.txt
OUTFILE2=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_sorted.bam
OUTBQSRTABLE=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_post_bqsr2.table
OUTFILE3=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_post_bqsr2.bam
OUTFILE4=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}.g.vcf

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk MarkDuplicates \
    --java-options "-Xmx64G"\
    -I $FILE\
    -O $OUTFILE\
    -M $MARKFILE\
    -CREATE_INDEX true
    
singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk SortSam \
    --java-options "-Xmx64G"\
    -I $OUTFILE\
    -O $OUTFILE2\
    -CREATE_INDEX true\
    -SO coordinate

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk BaseRecalibrator \
    --java-options "-Xmx64G"\
    -I $OUTFILE2\
    -O $OUTBQSRTABLE\
    -known-sites $known_sites\
    -R $ref\
    -L $BED

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk ApplyBQSR \
    --java-options "-Xmx64G"\
    -bqsr-recal-file $OUTBQSRTABLE\
    -I $OUTFILE2 \
    -O $OUTFILE3 \
    -R $ref\
    -L $BED

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk HaplotypeCaller \
    --java-options "-Xmx64G"\
    -I $OUTFILE3 \
    -O $OUTFILE4 \
    -R $ref\
    -native-pair-hmm-threads 20\
    -L $BED\
    -ERC GVCF\
    -G StandardAnnotation\
    -G AS_StandardAnnotation

```

```{bash bqsrplot}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-47
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/bqsrplot_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/bqsrplot_%A_%a.errcd
#SBATCH --job-name=caller
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main


cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp.sh

ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa
known_sites=/project/wiemels_260/sebastian/dependencies/gatk/refs/All_20180418-3.vcf.gz
BED=/project/wiemels_260/sebastian/dependencies/gatk/refs/ace_liftover_r5.bed

INPUTBAM=(`ls /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/*_post_bqsr2.bam | sed -n ${SLURM_ARRAY_TASK_ID}p`)
echo ${INPUTBAM}

base=$(basename ${INPUTBAM} _post_bqsr2.bam)

OUTBQSRTABLE2=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_after_bqsr2.table
OUTBQSRTABLE=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_post_bqsr2.table

OUTCSV=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_AnalyzeCovariates.csv
OUTPDF=/project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/${base}_AnalyzeCovariates.pdf

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk BaseRecalibrator \
    --java-options "-Xmx64G"\
    -I $INPUTBAM \
    -O $OUTBQSRTABLE2 \
    -known-sites $known_sites\
    -R $ref\
    -L $BED

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk AnalyzeCovariates \
    --java-options "-Xmx64G"\
    -before $OUTBQSRTABLE\
    -after $OUTBQSRTABLE2\
    -plots $OUTPDF\
    -csv $OUTCSV

```

```{bash dbimport}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=END
#SBATCH --array=1-25
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/dbimp_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/dbimp_%A_%a.errcd
#SBATCH --job-name=dbimp
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller
#sbatch /project/lishaobo_902/sebastian/rhabdo/codes/temp2.sh

# create a sample map file, rhabdo.sample.map
## to create a tab in terminal, do echo -e '[\t]'
## tab-delimited text file with sample_name--tab--path_to_sample_vcf per line
#sample1       sample1.vcf.gz


POS=(`cat /project/wiemels_260/sebastian/dependencies/gatk/refs/chrom_num.text | sed -n ${SLURM_ARRAY_TASK_ID}p`)
ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa

mkdir -p /project/lishaobo_902/sebastian/rhabdo/snp_dis/dbimport/${POS}
POS_dir=/project/lishaobo_902/sebastian/rhabdo/snp_dis/dbimport/${POS}

OUTFILE=${POS_dir}/${POS}.vcf.gz

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk GenomicsDBImport\
    --java-options "-Xmx64G"\
    -genomicsdb-workspace-path $POS_dir\
    -L $POS\
    -sample-name-map rhabdo.sample.map\
    -reader-threads 20\
    -batch-size 1

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif \
  gatk GenotypeGVCFs\
    --java-options "-Xmx64G"\
    -R $ref \
    -V gendb://$POS_dir\
    -G StandardAnnotation \
    -new-qual true \
    -O $OUTFILE


```

```{bash merge vcfs}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=ALL
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/mergevcfs_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/mergevcfs_%A_%a.errcd
#SBATCH --job-name=mergevcfs
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main


cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller

# create a file of MergeVcfs.list containing all paths of vcf.gz files
##/project/lishaobo_902/sebastian/rhabdo/snp_dis/dbimport/chr1/chr1.vcf.gz
##/project/lishaobo_902/sebastian/rhabdo/snp_dis/dbimport/chr2/chr2.vcf.gz
##...

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk MergeVcfs\
    --java-options "-Xmx64G"\
    -I /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/MergeVcfs.list\
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/chr.all.vcf.gz


```

filtering SNP calling

```{bash hard filtering}

#!/bin/bash
#SBATCH --time=15:00:00
#SBATCH --mail-type=ALL
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/vqsr_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/vqsr_%A_%a.errcd
#SBATCH --job-name=vqsr
#SBATCH -c 20
#SBATCH --ntasks=1
#SBATCH --mem=60G
#SBATCH --partition=main

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller
ref=/project/wiemels_260/sebastian/dependencies/gatk/refs/GRCh38_full_analysis_set_plus_decoy_hla.fa

bcftools view chr.all.vcf.gz  -i  "FMT/DP>8 & FMT/GQ>20" \
  -o chr.all.dg.vcf.gz \
  -O z --threads 20
  
bcftools view chr.all.dg.vcf.gz -i "AVG(FORMAT/GQ)>35 & F_MISSING < 0.1" \
  -o chr.all.avg.gq.vcf.gz \
  -O z --threads 20

tabix chr.all.avg.gq.vcf.gz

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk VariantRecalibrator\
    --java-options "-Xmx50G"\
    -R $ref \
    -V /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/chr.all.avg.gq.vcf.gz \
    --resource:hapmap,known=false,training=true,truth=true,prior=15.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/hapmap_3.3.hg38.vcf.gz \
    --resource:omni,known=false,training=true,truth=false,prior=12.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/1000G_omni2.5.hg38.vcf.gz \
    --resource:1000G,known=false,training=true,truth=false,prior=10.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/1000G_phase1.snps.high_confidence.hg38.vcf.gz \
    --resource:dbsnp,known=true,training=false,truth=false,prior=2.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/Homo_sapiens_assembly38.dbsnp138.vcf \
    -an QD -an MQ -an MQRankSum -an ReadPosRankSum -an FS -an SOR \
    -mode SNP \
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_SNP.recal\
    --tranches-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_SNP.tranches \
    --rscript-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_SNP.plots.R

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk ApplyVQSR\
    --java-options "-Xmx50G"\
    -R $ref \
    -V /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/chr.all.avg.gq.vcf.gz \
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.snp.vcf.gz \
    --truth-sensitivity-filter-level 99.0 \
    --tranches-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_SNP.tranches \
    --recal-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_SNP.recal \
    -mode SNP

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk VariantRecalibrator\
    --java-options "-Xmx50G"\
    -R $ref \
    -V /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/chr.all.avg.gq.vcf.gz \
    --max-gaussians 4 \
    --resource:mills,known=false,training=true,truth=true,prior=12.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/Mills_and_1000G_gold_standard.indels.hg38.vcf.gz \
    --resource:dbsnp,known=true,training=false,truth=false,prior=2.0 /project/desmith_488/sebastian/gatk/refs/GATK_resource_bundle/Homo_sapiens_assembly38.dbsnp138.vcf \
    -an QD -an MQ -an MQRankSum -an ReadPosRankSum -an FS -an SOR -an InbreedingCoeff \
    -mode INDEL \
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_INDEL.recal\
    --tranches-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_INDEL.tranches \
    --rscript-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_INDEL.plots.R

singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk ApplyVQSR\
    --java-options "-Xmx50G"\
    -R $ref \
    -V /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/chr.all.avg.gq.vcf.gz \
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.indel.vcf.gz \
    --truth-sensitivity-filter-level 99.0 \
    --tranches-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_INDEL.tranches \
    --recal-file /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/output_INDEL.recal \
    -mode INDEL

#singularity run \
#  --bind /project/lishaobo_902/sebastian/rhabdo/snp_dis:/snp_dis \
#  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif
  
singularity exec \
  --bind /project:/project\
  /project/wiemels_260/sebastian/dependencies/gatk/gatk_latest.sif\
  gatk MergeVcfs\
    --java-options "-Xmx64G"\
    -I /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.snp.vcf.gz\
    -I /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.indel.vcf.gz\
    -O /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.vcf.gz

bgzip -d /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.vcf.gz

```

filtering SNP calling with CNN 

```{bash}

   
```



```{bash annotate}

#!/bin/bash
#SBATCH --time=24:00:00
#SBATCH --mail-type=ALL
#SBATCH --job-name=anno
#SBATCH --output=/project/lishaobo_902/sebastian/rhabdo/snp_dis/output/anno_%A_%a.out
#SBATCH --error=/project/lishaobo_902/sebastian/rhabdo/snp_dis/error/anno_%A_%a.errcd
#SBATCH --mem=45G
#SBATCH -c 20
#SBATCH --ntasks=1

module load gcc/11.3.0 htslib/1.14

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/annovar

WGSNV=/project/desmith_488/sebastian/gatk/refs/snp_databases/cadd/whole_genome_SNVs.tsv.gz
INDELS=/project/desmith_488/sebastian/gatk/refs/snp_databases/cadd/InDels.tsv.gz
CADD=/project/desmith_488/sebastian/gatk/refs/snp_databases/cadd/cadd.hdr
TOPMED=/project/desmith_488/sebastian/gatk/refs/snp_databases/topmed/bravo-dbsnp-all.tsv.gz

#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar refGene humandb/
#annotate_variation.pl -buildver hg38 -downdb cytoBand humandb/
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar exac03 humandb/ 
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar avsnp150 humandb/ 
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar dbnsfp42c humandb/
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar gnomad211_genome humandb/
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar gnomad211_exome humandb/
#annotate_variation.pl -buildver hg38 -downdb -webfrom annovar clinvar_20210501 humandb/

table_annovar.pl /project/lishaobo_902/sebastian/rhabdo/snp_dis/HaplotypeCaller/rhabdo.calling.vcf \
  /project/wiemels_260/sebastian/dependencies/annovar/annovar/humandb/ \
  -buildver hg38 \
  -out annotated \
  -remove \
  -protocol refGene,cytoBand,exac03,avsnp150,dbnsfp42c,gnomad211_genome,gnomad211_exome,clinvar_20210501\
  -operation g,r,f,f,f,f,f,f \
  -vcfinput \
  --dot2underline \
  --polish -xref /project/wiemels_260/sebastian/dependencies/annovar/annovar/example/gene_fullxref.txt \
  --thread 20 \
  --nastring .

awk '{gsub(/^chr/,""); print}' annotated.hg38_multianno.vcf > annotated.rhabdo.calling.no.chr.vcf
bgzip annotated.rhabdo.calling.no.chr.vcf
tabix annotated.rhabdo.calling.no.chr.vcf.gz

bcftools annotate -a $INDELS -h $CADD -c Chrom,Pos,Ref,Alt,RawScore,PHRED annotated.rhabdo.calling.no.chr.vcf.gz -O z -o annotated.rhabdo.hg38_cadd_indel_output.vcf.gz --threads 20

bcftools annotate -a $WGSNV -h $CADD -c Chrom,Pos,Ref,Alt,RawScore,PHRED annotated.rhabdo.hg38_cadd_indel_output.vcf.gz -O z -o annotated.rhabdo.hg38_cadd_wg_output.vcf.gz --threads 20
tabix annotated.rhabdo.hg38_cadd_wg_output.vcf.gz
bcftools annotate -a $TOPMED -h $CADD -c Chrom,Pos,Ref,Alt,TOPMED annotated.rhabdo.hg38_cadd_wg_output.vcf.gz -O z -o annotated.rhabdo.hg38.vcf.gz --threads 20
tabix annotated.rhabdo.hg38.vcf.gz


```

```{r check out what SNPs were identified}

setwd("/project/lishaobo_902/sebastian/rhabdo/snp_dis")
library(tidyverse)
library(data.table)
library(vcfR)

vcfR = read.vcfR("annovar/annotated.rhabdo.hg38.vcf.gz")

vcf_data = vcfR2tidy(vcfR, dot_is_NA=TRUE, single_frame=TRUE)$dat %>% 
  as.data.frame

snp_summary = vcf_data %>%
  mutate(ID = paste0(CHROM,":",POS,":",REF,":",ALT)) %>%
  dplyr::select(ID,CHROM,POS,REF,ALT,Gene_refGene,gt_GT) %>%
  separate(gt_GT,c("ale1","ale2")) %>%
  mutate(mac = as.numeric(ale1)+as.numeric(ale2)) %>%
  dplyr::select(-ale1,-ale2) %>%
  group_by(ID,CHROM,POS,REF,ALT,Gene_refGene)

snp_summarized = snp_summary %>%
  summarize(mac_snp = sum(mac),
            maf_snp = mac_snp/(n()*2)) 
  
snp_summarized_df = snp_summarized %>%
  as.data.frame()

anno_input = snp_summarized_df %>%
  mutate(POS2=POS) %>%
  arrange(CHROM,POS) %>%
  dplyr::select(CHROM,POS,POS2,ALT,REF,ID)

anno_input2 = snp_summarized_df %>%
  mutate(POS2=POS) %>%
  arrange(CHROM,POS) %>%
  dplyr::select(CHROM,POS,POS2,REF,ALT,ID)
  
write.csv(snp_summarized_df,"all.snps.csv")

write.table(anno_input,"plinks/rhabdo.hg38.input.anno.alt.ref.txt",
            sep='\t',
            quote=FALSE,
            row.names = FALSE,
            col.names = FALSE)

write.table(anno_input2,"plinks/rhabdo.hg38.input.anno.ref.alt.txt",
            sep='\t',
            quote=FALSE,
            row.names = FALSE,
            col.names = FALSE)

```


```{bash convert output to PLINK and annotate with rs ID and MAF first}

# bi-alleles only 
cd /project/lishaobo_902/sebastian/rhabdo/snp_dis/plinks

#annotate_variation.pl -out anno -build hg38 rhabdo.hg38.input.anno.txt /project/wiemels_260/sebastian/dependencies/annovar/annovar/humandb/
annotate_variation.pl rhabdo.hg38.input.anno.alt.ref.txt -build hg38 -filter -dbtype avsnp150 /project/wiemels_260/sebastian/dependencies/annovar/annovar/humandb/ 
annotate_variation.pl rhabdo.hg38.input.anno.ref.alt.txt -build hg38 -filter -dbtype avsnp150 /project/wiemels_260/sebastian/dependencies/annovar/annovar/humandb/ 

# 170204 if ref alt
# 1828 if alt ref

```

```{r combine all the generated files}

setwd("/project/lishaobo_902/sebastian/rhabdo/snp_dis/plinks")
library(tidyverse)
library(data.table)

snps1 = fread("rhabdo.hg38.input.anno.ref.alt.txt.hg38_avsnp150_dropped")%>%
  as.data.frame() %>%
  dplyr::mutate(ID = V8,
                rsId = V2,
                CHROM = V3,
                POS = V4) 

snps2 = fread("rhabdo.hg38.input.anno.alt.ref.txt.hg38_avsnp150_dropped")%>%
  as.data.frame() %>%
  dplyr::mutate(ID = V8,
                rsId = V2,
                CHROM = V3,
                POS = V4) 

snps = rbind(snps1,snps2) %>%
  distinct(ID,.keep_all=TRUE) %>%
  dplyr::select(rsId,CHROM,POS)

snp_summarized_df = read.csv("../all.snps.csv") %>%
  as.data.frame() %>%
  dplyr::select(-X) %>%
  arrange(CHROM,POS) %>%
  dplyr::left_join(snps,by=c("CHROM","POS")) %>%
  mutate(gene_name = str_replace_all(Gene_refGene,"\\\\x",";")) %>%
  dplyr::select(-mac_snp,-Gene_refGene)

write.table(snp_summarized_df,"/project/lishaobo_902/sebastian/rhabdo/snp_dis/all.snps.annotated.txt",
            sep='\t',
            quote=FALSE,
            row.names = FALSE)

snp_df = read.csv("../all.snps.csv",row.names = 1) %>%
  as.data.frame()

write.table(snp_df,"../all.snps.txt",
            sep='\t',
            quote=FALSE,
            row.names = FALSE)

```

look up analysis

```{bash}

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis
# PMID: 34095712
# collaborator in Chicago: CBFB, RUNX1, RUNX2, RUNX3, NOTCH1
# doi: https://doi.org/10.1101/2022.03.01.22271578

#a1 = a %>% dplyr::distinct(V1,.keep_all=TRUE) %>% arrange(V1)
#write.table(a1,"a1.txt",sep='\t',row.names=FALSE,col.names=FALSE,quote=FALSE)

# 325 genes (TNFRSF6 (FAS) are split into 2)
grep -wf gene.list.cancer.txt all.snps.txt > all.snps.selected.txt
# 4576 / 188947

```

global and local ancestries of subjects

```{bash Local ancestry regression in latino ancestry subjects}
#Local ancestry regression in latino ancestry subjects using RFmix

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis

for chr in {1..22};do
  java -jar /project/wiemels_260/sebastian/dependencies/beagle5/conform-gt.24May16.cee.jar\
    gt=/project/lishaobo_902/sebastian/rhabdo/snp_dis/annovar/annotated.rhabdo.hg38.vcf.gz\
    ref=/project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.shapeit2_integrated_snvindels_v2a_27022019.GRCh38.phased.vcf.gz\
    chrom=$chr\
    match=POS\
    out=/scratch2/lishaobo/rhabdo/$chr.rhabdo.hg38
  
  tabix /scratch2/lishaobo/rhabdo/$chr.rhabdo.hg38.vcf.gz
  
done    

## converting ref vcf.gz file to bref3 files for faster phasing
#for chr in {1..22};do
#  java -jar /project/wiemels_260/sebastian/dependencies/beagle5/bref3.18May20.d20.jar\
#    /project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.shapeit2_integrated_snvindels_v2a_27022019.GRCh38.phased.vcf.gz > #/project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.bref3
#done

## phase vcf files with beagle5
for chr in {1..22};do
  java -jar /project/wiemels_260/sebastian/dependencies/beagle5/beagle.18May20.d20.jar\
      gt=/scratch2/lishaobo/rhabdo/$chr.rhabdo.hg38.vcf.gz\
      ref=/project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.bref3\
      out=/scratch2/lishaobo/rhabdo/$chr.rhabdo.hg38.query.phased\
      map=/project/wiemels_260/sebastian/dependencies/beagle5/plink.GRCh38.map/plink.chr$chr.GRCh38.map
  
  tabix /scratch2/lishaobo/rhabdo/$chr.rhabdo.hg38.query.phased.vcf.gz
    
done  

# count the number of SNPs per reviewer question
#for chr in {1..22};do
#  echo chr$chr
#  bcftools stats $chr.query.phased.vcf.gz | grep "number of SNPs:"
#done

######################
# A phased VCF/BCF file containing reference haplotypes (in any order)
# ref panel has too many SNPs. Only include overlaps with lat

cd /project/lishaobo_902/sebastian/rhabdo/snp_dis

# sample_id90.txt includes five ancestries with the purest genetics

plink --bfile /project/wiemels_494/sebastian/temp_scratch_backup/populationRisk2/hgdp_1kg\
  --recode vcf bgz\
  --keep sample_id90.txt\
  --out /scratch2/lishaobo/rhabdo/panel.ref
# 450 subjects, 90 from each ancestry

## make reference data conforming to that of reference data
for chr in {1..22};do
  java -jar /project/wiemels_260/sebastian/dependencies/beagle5/conform-gt.24May16.cee.jar\
    gt=/scratch2/lishaobo/rhabdo/panel.ref.vcf.gz\
    ref=/project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.shapeit2_integrated_snvindels_v2a_27022019.GRCh38.phased.vcf.gz\
    chrom=$chr\
    match=POS\
    out=/scratch2/lishaobo/rhabdo/$chr.panel.ref
    
    tabix /scratch2/lishaobo/rhabdo/$chr.panel.ref.vcf.gz

done    

## phase vcf files with beagle5
for chr in {1..22};do
  java -Xmx100g -jar /project/wiemels_260/sebastian/dependencies/beagle5/beagle.18May20.d20.jar\
    gt=/scratch2/lishaobo/rhabdo/$chr.panel.ref.vcf.gz\
    ref=/project/wiemels_260/sebastian/gwas_1000G/GRCh38/ALL.chr$chr.bref3\
    out=/scratch2/lishaobo/rhabdo/$chr.panel.phased\
    map=/project/wiemels_260/sebastian/dependencies/beagle5/plink.GRCh38.map/plink.chr$chr.GRCh38.map

  tabix /scratch2/lishaobo/rhabdo/$chr.panel.phased.vcf.gz

done  

#for chr in {1..22};do
#  echo chr$chr
#  bcftools stats $chr.panel.phased.vcf.gz | grep "number of SNPs:"
#done

######################
# Manipulating map file for rfmix
#chr=1
#awk -v OFS='\t' '{print $1,$4,$3}' /project/wiemels_260/sebastian/dependencies/beagle5/plink.GRCh38.map/plink.chr$chr.GRCh38.map > #/project/wiemels_260/sebastian/dependencies/rfmix/rfmix.GRCh38.map/rfmix.GRCh38.map
#for chr in {2..22};do
#  awk -v OFS='\t' '{print $1,$4,$3}' /project/wiemels_260/sebastian/dependencies/beagle5/plink.GRCh38.map/plink.chr$chr.GRCh38.map >> #/project/wiemels_260/sebastian/dependencies/rfmix/rfmix.GRCh38.map/rfmix.GRCh38.map
#done

```

```{bash script for rfmix, bcf file}
#!/bin/bash
#SBATCH --time=48:00:00
#SBATCH --mail-type=END
#SBATCH --job-name=rgnbcf
#SBATCH --array=1-22
#SBATCH --mem=160G

#sbatch /project/wiemels_260/sebastian/gwas_glioma/codes/temp3.sh

cd /scratch2/lishaobo/rhabdo/

# creating bcf files for rfmix running

bcftools view -o ${SLURM_ARRAY_TASK_ID}.query.bcf -Ob\
  --threads 200\
  ${SLURM_ARRAY_TASK_ID}.rhabdo.hg38.query.phased.vcf.gz
bcftools index -f ${SLURM_ARRAY_TASK_ID}.query.bcf

bcftools view -o ${SLURM_ARRAY_TASK_ID}.panel.bcf -Ob\
  --threads 200\
  ${SLURM_ARRAY_TASK_ID}.panel.phased.vcf.gz
bcftools index -f ${SLURM_ARRAY_TASK_ID}.panel.bcf

rfmix\
    -f ${SLURM_ARRAY_TASK_ID}.query.bcf\
  	-r ${SLURM_ARRAY_TASK_ID}.panel.bcf\
  	-m /project/lishaobo_902/sebastian/rhabdo/snp_dis/sample_id90.txt\
  	-g /project/wiemels_260/sebastian/dependencies/rfmix/rfmix.GRCh38.map/rfmix.GRCh38.map\
  	-o ${SLURM_ARRAY_TASK_ID}.bcf.rgn.rhabdo\
  	--chromosome=${SLURM_ARRAY_TASK_ID}

```



