---
output: html_document
editor_options: 
  chunk_output_type: console
---

Project description: to investigate the associations between DNA methylation and birth order, combining datasets from multiple cohorts from 12 countries.

This manuscript is in preparation.

```{r updating files}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

neo = read.csv("sourceF/meth_bthdata_BC.csv") %>%
  as.data.frame()%>%
  dplyr::mutate(LiveBirthsNowLiving=as.numeric(as.character(LiveBirthsNowLiving)),
         TerminationsBefore20Weeks=as.numeric(as.character(TerminationsBefore20Weeks)),
         Terminations20WeeksPlus=as.numeric(as.character(Terminations20WeeksPlus)),
         parity = as.numeric(as.character(parity))) %>%
  dplyr::mutate(par_neo = LiveBirthsNowLiving+1,
                mo_gravid_neo = LiveBirthsNowLiving+
                  TerminationsBefore20Weeks+
                  Terminations20WeeksPlus+1,
                birth_order = 0,
                preg_order = 0)

for( i in 1:nrow(neo)){
  if(is.na(neo$parity[i])){
    neo$birth_order[i] = neo$par_neo[i]
    neo$preg_order[i] = neo$mo_gravid_neo[i]
  }else{
    neo$birth_order[i] = as.numeric(as.character(neo$parity))[i]
    neo$preg_order[i] = as.numeric(as.character(neo$mo_gravid_int))[i]
  }}

write.csv(neo,"sourceF/bthorder_meth_ph123_neo.csv")

```
# birthOrder analysis without genetic correction
```{r ewas of birthorder}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

# load birthorder data 

for(var in c("birth_order","preg_order")){
  
    # var = "birth_order"
  
  for(set in 1:4){
  
    #set=1
  
   if(set%in%c(1,2)){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       inner_join(bo, by="beadPosition")
     
     }
    
   if(set==3){
    
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
         
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
       filter(Trisomy21==0)%>%
       inner_join(bo, by="subjectId")
     
     }
    
   if(set==4){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       mutate(birth_order = as.numeric(as.character(parity)), 
              preg_order = as.numeric(as.character(mo_gravid_int))) %>%
       select(beadPosition, birth_order, preg_order)
  
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       filter(smp_type=="bg")%>%
       inner_join(bo, by="beadPosition")
    
     }
  
    clinical1 = clinical0 %>% 
      mutate(plate = as.factor(plate),
             sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,one_of(var)) %>%
      na.omit()
    
    rfs = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    genePC = read.table(paste0("~/mQTL/ccls.set",set,".pc.eigenvec"))
    colnames(genePC) = c("FID","subjectId","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10")
  
    clinical2 = genePC %>%
      inner_join(rfs, by="subjectId")
  
  
  # load methylation data
    
    methy = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy) = c()
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
  
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
    
    print(paste0("subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  

    probes = colnames(methyDF)
  
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$rc1 + 
              cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + cF$rc7 + cF$rc8 +
              cF$rc9 + cF$rc10 + cF$PC1 + cF$PC2 + cF$PC3 + cF$PC4 + cF$PC5 +
              cF$PC6 + cF$PC7 + cF$PC8 + cF$PC8 + cF$PC9 + cF$PC10
           ))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  
  
  }
}

```

```{bash}

cd ~/ewas/birthorder

sed -i 's/Std. Error/SE/g' *.ewas.txt

### 450K birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.birth_order.ewas.txt
PROCESS set2.birth_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.450K.birth_order.ewas.txt

### EPIC birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set3.birth_order.ewas.txt
PROCESS set4.birth_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.EPIC.birth_order.ewas.txt

### ALL birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.birth_order.ewas.txt
PROCESS set2.birth_order.ewas.txt
PROCESS set3.birth_order.ewas.txt
PROCESS set4.birth_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.all.birth_order.ewas.txt

### 450K preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.preg_order.ewas.txt
PROCESS set2.preg_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.450K.preg_order.ewas.txt

### EPIC preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set3.preg_order.ewas.txt
PROCESS set4.preg_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.EPIC.preg_order.ewas.txt

### ALL preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.preg_order.ewas.txt
PROCESS set2.preg_order.ewas.txt
PROCESS set3.preg_order.ewas.txt
PROCESS set4.preg_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.all.preg_order.ewas.txt

```

```{r annotation}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(qvalue)

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")
anno       <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC       <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

for(platform in c("450K","EPIC","all")){
  for(birth in c("preg_order","birth_order")){

  meta = fread(paste0("meta.",platform,".",birth,".ewas.txt")) %>%
    select(-Allele1, -Allele2) %>%
    mutate(cpg = gsub("_.*$","",MarkerName),
           snp = gsub("^.*_","",MarkerName)) %>%
    arrange(by=`P-value`) %>%
    mutate(qvalue = qvalue(`P-value`)$qvalues,
           bonferroni = p.adjust(`P-value`, method = "bonferroni", n = length(`P-value`)),
           fdr = p.adjust(`P-value`, method = "fdr", n = length(`P-value`))) %>%
    filter(fdr < 0.05) %>%
    left_join(anno,by="cpg") 

  print(paste0(nrow(meta)," probes are significant for platform ",platform," for variable ",birth))

  if(nrow(meta)>0) write.csv(meta,paste0(birth,".",platform,".sig.meta.csv"))
  
    }}

```

[1] "0 probes are significant for platform 450K for variable preg_order"
[1] "0 probes are significant for platform 450K for variable birth_order"
[1] "1 probes are significant for platform EPIC for variable preg_order"
[1] "0 probes are significant for platform EPIC for variable birth_order"
[1] "1 probes are significant for platform all for variable preg_order"
[1] "4 probes are significant for platform all for variable birth_order"

```{r plotting cg24567724 with birth order}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

cg = "cg24567724"

for(set in c(3,4)){
  for(birth in c("preg_order","birth_order")){

    setN = paste0("set",set)
    
    if(set==3){
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
    
    clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
     filter(Trisomy21==0)%>%
     inner_join(bo, by="subjectId") %>%
     select(beadPosition,one_of(birth))
      
    }

    if(set==4){
   bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123.csv") %>%
     filter(set == setN)%>%
      mutate(birth_order = as.numeric(as.character(parity)), 
      preg_order = as.numeric(as.character(mo_gravid_int))) %>%
      na.omit() %>%
      select(beadPosition, one_of(birth))
   
   clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
     filter(smp_type=="bg")%>%
     inner_join(bo, by="beadPosition")%>%
     select(beadPosition,one_of(birth))
    }
    

    methy = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds")) 

    methy1 = methy[methy$probeId == cg,] 
    rownames(methy1) <- c()
    
    methy2 = methy1 %>%
      column_to_rownames(var="probeId") %>%
      t %>%
      as.data.frame() %>%
      na.omit()%>%
      rownames_to_column(var="beadPosition") %>%
      inner_join(clinical0,by="beadPosition") 
    
    methy2[,birth] = as.factor(methy2[,birth])
      
    
    methy2 %>%
      ggplot(aes_string(x=birth,y=cg)) +
      geom_point()+
      geom_smooth(method=lm)+
      theme_bw()+
      labs(x=birth,y=paste(cg,"methylation"))+
      ggtitle(paste("Correlation between",cg,"methylation and",birth))+
      theme(plot.title = element_text(hjust = 0.5))
    ggsave(paste0("set",set,".",birth,".",cg,".png"),width = 8, height = 6)

     methy2 %>%
      ggplot(aes_string(x=birth,y=cg)) +
      geom_violin(trim=FALSE)+
      theme_bw()+
      labs(x=birth,y=paste(cg,"methylation"))+
      ggtitle(paste("Correlation between",cg,"methylation and",birth))+
      theme(plot.title = element_text(hjust = 0.5))
    ggsave(paste0("set",set,".",birth,".",cg,".violin.png"),width = 8, height = 6)


  }}


```

# birthOrder analysis correcting for birth related variables, without genetic correction

```{r ewas of birthorder, correcting for vars}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

# load birthorder data 

for(var in c("birth_order","preg_order")){
  
  # var = "birth_order"

  for(set in 1:4){
  
    #set=1
  
   if(set%in%c(1,2)){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       inner_join(bo, by="beadPosition")
     
     }
    
   if(set==3){
    
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
         
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
       filter(Trisomy21==0)%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=Y_birthweight,
              gestage=Y_gestation_week)%>%
       inner_join(bo, by="subjectId")
     
     }
    
   if(set==4){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       mutate(birth_order = as.numeric(as.character(parity)), 
              preg_order = as.numeric(as.character(mo_gravid_int))) %>%
       select(beadPosition, birth_order, preg_order)
  
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       filter(smp_type=="bg")%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=ch_birthwt_bc_full,
              gestage=ch_gestage_int)%>%
       inner_join(bo, by="beadPosition")
    
     }
  
    clinical1 = clinical0 %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,birthWt,gestage,one_of(var)) %>%
      na.omit()
    
    rfs = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    genePC = read.table(paste0("~/mQTL/ccls.set",set,".pc.eigenvec"))
    colnames(genePC) = c("FID","subjectId","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10")
  
    clinical2 = genePC %>%
      inner_join(rfs, by="subjectId")
  
  
  # load methylation data
    
    methy = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy) = c()
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
  
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
      
    probes = colnames(methyDF)
  
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$gestage + cF$birthWt + cF$rc1 + 
              cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + cF$rc7 + cF$rc8 +
              cF$rc9 + cF$rc10 + cF$PC1 + cF$PC2 + cF$PC3 + cF$PC4 + cF$PC5 +
              cF$PC6 + cF$PC7 + cF$PC8 + cF$PC8 + cF$PC9 + cF$PC10
           ))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("set",set,".",var,".vars.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

  }
}

```

```{bash}

cd ~/ewas/birthorder

sed -i 's/Std. Error/SE/g' *.vars.ewas.txt

### 450K birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.birth_order.vars.ewas.txt
PROCESS set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.450K.birth_order.vars.ewas.txt

### EPIC birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set3.birth_order.vars.ewas.txt
PROCESS set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.EPIC.birth_order.vars.ewas.txt

### ALL birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.birth_order.vars.ewas.txt
PROCESS set2.birth_order.vars.ewas.txt
PROCESS set3.birth_order.vars.ewas.txt
PROCESS set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.all.birth_order.vars.ewas.txt

### 450K preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.preg_order.vars.ewas.txt
PROCESS set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.450K.preg_order.vars.ewas.txt

### EPIC preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set3.preg_order.vars.ewas.txt
PROCESS set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.EPIC.preg_order.vars.ewas.txt

### ALL preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS set1.preg_order.vars.ewas.txt
PROCESS set2.preg_order.vars.ewas.txt
PROCESS set3.preg_order.vars.ewas.txt
PROCESS set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL meta.all.preg_order.vars.ewas.txt

```

meta.450K.birth_order.vars: Smallest p-value is 1.87e-06 at marker 'cg01396391'
meta.EPIC.birth_order.vars: Smallest p-value is 6.402e-07 at marker 'cg03536404'
meta.all.birth_order.vars: Smallest p-value is 8.48e-08 at marker 'cg27293155'

```{r annotation}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(qvalue)

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")
anno       <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC       <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

for(platform in c("450K","EPIC","all")){
  for(birth in c("preg_order","birth_order")){

  meta = fread(paste0("meta.",platform,".",birth,".vars.ewas.txt")) %>%
    select(-Allele1, -Allele2) %>%
    mutate(cpg = gsub("_.*$","",MarkerName),
           snp = gsub("^.*_","",MarkerName)) %>%
    arrange(by=`P-value`) %>%
    mutate(qvalue = qvalue(`P-value`)$qvalues,
           bonferroni = p.adjust(`P-value`, method = "bonferroni", n = length(`P-value`)),
           fdr = p.adjust(`P-value`, method = "fdr", n = length(`P-value`))) %>%
    filter(fdr < 0.05) %>%
    left_join(anno,by="cpg") 

  print(paste0(nrow(meta)," probes are significant for platform ",platform," for variable ",birth))

  if(nrow(meta)>0) write.csv(meta,paste0(birth,".",platform,".sig.vars.meta.csv"))
  
    }}

```
[1] "0 probes are significant for platform 450K for variable preg_order"
[1] "0 probes are significant for platform 450K for variable birth_order"
[1] "0 probes are significant for platform EPIC for variable preg_order"
[1] "0 probes are significant for platform EPIC for variable birth_order"
[1] "0 probes are significant for platform all for variable preg_order"
[1] "0 probes are significant for platform all for variable birth_order"
# birthOrder analysis correcting for birth related variables, by ancestry
```{r ewas of birthorder, correcting for vars}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

# load birthorder data 

for(var in c("birth_order","preg_order")){
  
  # var = "birth_order"

  for(set in 1:4){
  
    #set=1
  
   if(set%in%c(1,2)){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       inner_join(bo, by="beadPosition") 
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==3){
    
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
         
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
       filter(Trisomy21==0)%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=Y_birthweight,
              gestage=Y_gestation_week)%>%
       inner_join(bo, by="subjectId")
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==4){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       mutate(birth_order = as.numeric(as.character(parity)), 
              preg_order = as.numeric(as.character(mo_gravid_int))) %>%
       select(beadPosition, birth_order, preg_order)
  
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       filter(smp_type=="bg")%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=ch_birthwt_bc_full,
              gestage=ch_gestage_int)%>%
       inner_join(bo, by="beadPosition")

     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
    
     }
  
    
    # load methylation data
    
    methy = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy) = c()
    
    # EUR  
    race="eur"
    
    clinical1 = clinical_eur %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,birthWt,gestage,one_of(var)) %>%
      na.omit()
    
    rfs = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    genePC = read.table(paste0("~/mQTL/ccls.set",set,".pc.eigenvec"))
    colnames(genePC) = c("FID","subjectId","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10")
  
    clinical2 = genePC %>%
      inner_join(rfs, by="subjectId")
  
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
  
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]

    print(paste0("subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
      
    probes = colnames(methyDF)
  
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$gestage + cF$birthWt + cF$rc1 + 
              cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + cF$rc7 + cF$rc8 +
              cF$rc9 + cF$rc10 + cF$PC1 + cF$PC2 + cF$PC3 + cF$PC4 + cF$PC5 +
              cF$PC6 + cF$PC7 + cF$PC8 + cF$PC8 + cF$PC9 + cF$PC10
           ))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0(race,".set",set,".",var,".vars.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  
    
    
    # LAT  
    race="lat"
    
    clinical1 = clinical_lat %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,birthWt,gestage,one_of(var)) %>%
      na.omit()
    
    rfs = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    genePC = read.table(paste0("~/mQTL/ccls.set",set,".pc.eigenvec"))
    colnames(genePC) = c("FID","subjectId","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","PC9","PC10")
    
    clinical2 = genePC %>%
      inner_join(rfs, by="subjectId")
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
    
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
      
    print(paste0("subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
    
    probes = colnames(methyDF)
    
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$gestage + cF$birthWt + cF$rc1 + 
              cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + cF$rc7 + cF$rc8 +
              cF$rc9 + cF$rc10 + cF$PC1 + cF$PC2 + cF$PC3 + cF$PC4 + cF$PC5 +
              cF$PC6 + cF$PC7 + cF$PC8 + cF$PC8 + cF$PC9 + cF$PC10
           ))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0(race,".set",set,".",var,".vars.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

  }
}

```

```{bash}

cd ~/ewas/birthorder

sed -i 's/Std. Error/SE/g' *.vars.ewas.txt

### 450K birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.birth_order.vars.ewas.txt
PROCESS eur.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.450K.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set1.birth_order.vars.ewas.txt
PROCESS lat.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.450K.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.birth_order.vars.ewas.txt
PROCESS eur.set2.birth_order.vars.ewas.txt
PROCESS lat.set1.birth_order.vars.ewas.txt
PROCESS lat.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.450K.birth_order.vars.ewas.txt

### EPIC birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set3.birth_order.vars.ewas.txt
PROCESS eur.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.EPIC.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set3.birth_order.vars.ewas.txt
PROCESS lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.EPIC.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set3.birth_order.vars.ewas.txt
PROCESS eur.set4.birth_order.vars.ewas.txt
PROCESS lat.set3.birth_order.vars.ewas.txt
PROCESS lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.EPIC.birth_order.vars.ewas.txt

### ALL birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.birth_order.vars.ewas.txt
PROCESS eur.set2.birth_order.vars.ewas.txt
PROCESS eur.set3.birth_order.vars.ewas.txt
PROCESS eur.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.all.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set1.birth_order.vars.ewas.txt
PROCESS lat.set2.birth_order.vars.ewas.txt
PROCESS lat.set3.birth_order.vars.ewas.txt
PROCESS lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.all.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.birth_order.vars.ewas.txt
PROCESS eur.set2.birth_order.vars.ewas.txt
PROCESS eur.set3.birth_order.vars.ewas.txt
PROCESS eur.set4.birth_order.vars.ewas.txt
PROCESS lat.set1.birth_order.vars.ewas.txt
PROCESS lat.set2.birth_order.vars.ewas.txt
PROCESS lat.set3.birth_order.vars.ewas.txt
PROCESS lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.all.birth_order.vars.ewas.txt



### 450K preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.preg_order.vars.ewas.txt
PROCESS eur.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.450K.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set1.preg_order.vars.ewas.txt
PROCESS lat.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.450K.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.preg_order.vars.ewas.txt
PROCESS eur.set2.preg_order.vars.ewas.txt
PROCESS lat.set1.preg_order.vars.ewas.txt
PROCESS lat.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.450K.preg_order.vars.ewas.txt

### EPIC preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set3.preg_order.vars.ewas.txt
PROCESS eur.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.EPIC.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set3.preg_order.vars.ewas.txt
PROCESS lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.EPIC.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set3.preg_order.vars.ewas.txt
PROCESS eur.set4.preg_order.vars.ewas.txt
PROCESS lat.set3.preg_order.vars.ewas.txt
PROCESS lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.EPIC.preg_order.vars.ewas.txt

### ALL preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.preg_order.vars.ewas.txt
PROCESS eur.set2.preg_order.vars.ewas.txt
PROCESS eur.set3.preg_order.vars.ewas.txt
PROCESS eur.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL eur.meta.all.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS lat.set1.preg_order.vars.ewas.txt
PROCESS lat.set2.preg_order.vars.ewas.txt
PROCESS lat.set3.preg_order.vars.ewas.txt
PROCESS lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL lat.meta.all.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS eur.set1.preg_order.vars.ewas.txt
PROCESS eur.set2.preg_order.vars.ewas.txt
PROCESS eur.set3.preg_order.vars.ewas.txt
PROCESS eur.set4.preg_order.vars.ewas.txt
PROCESS lat.set1.preg_order.vars.ewas.txt
PROCESS lat.set2.preg_order.vars.ewas.txt
PROCESS lat.set3.preg_order.vars.ewas.txt
PROCESS lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL byrace.meta.all.preg_order.vars.ewas.txt




```

```{r annotation}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(qvalue)

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")
anno       <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC       <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

for(platform in c("450K","EPIC","all")){
  for(birth in c("preg_order","birth_order")){
    for(race in c("eur","lat","byrace")){

  meta = fread(paste0(race,".meta.",platform,".",birth,".vars.ewas.txt")) %>%
    select(-Allele1, -Allele2) %>%
    mutate(cpg = gsub("_.*$","",MarkerName),
           snp = gsub("^.*_","",MarkerName)) %>%
    arrange(by=`P-value`) %>%
    mutate(qvalue = qvalue(`P-value`)$qvalues,
           bonferroni = p.adjust(`P-value`, method = "bonferroni", n = length(`P-value`)),
           fdr = p.adjust(`P-value`, method = "fdr", n = length(`P-value`))) %>%
    filter(fdr < 0.05) %>%
    left_join(anno,by="cpg") 

  print(paste0(nrow(meta)," probes are significant for platform ",platform," for variable ",birth))

  if(nrow(meta)>0) write.csv(meta,paste0(birth,".",platform,".sig.vars.meta.csv"))
  
    }}}

```

# birthOrder analysis correcting for birth related variables, by ancestry, no ancestry PCs

```{r ewas of birthorder, correcting for vars}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

# load birthorder data 

for(var in c("birth_order","preg_order")){
  
  # var = "birth_order"

  for(set in 1:4){
  
    #set=1
  
   if(set%in%c(1,2)){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       inner_join(bo, by="beadPosition") 
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==3){
    
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
         
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
       filter(Trisomy21==0)%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=Y_birthweight,
              gestage=Y_gestation_week)%>%
       inner_join(bo, by="subjectId")
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==4){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       mutate(birth_order = as.numeric(as.character(parity)), 
              preg_order = as.numeric(as.character(mo_gravid_int))) %>%
       select(beadPosition, birth_order, preg_order)
  
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       filter(smp_type=="bg")%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=ch_birthwt_bc_full,
              gestage=ch_gestage_int)%>%
       inner_join(bo, by="beadPosition")

     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
    
     }
  
    
    # load methylation data
    
    methy = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy) = c()
    
    # EUR  
    race="eur"
    
    clinical1 = clinical_eur %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,birthWt,gestage,one_of(var)) %>%
      na.omit()
    
    clinical2 = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
  
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]

    print(paste0("no PC, subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("no PC, probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
      
    probes = colnames(methyDF)
  
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$gestage + 
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".vars.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  
    
    
    # LAT  
    race="lat"
    
    clinical1 = clinical_lat %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,sex,plate,birthWt,gestage,one_of(var)) %>%
      na.omit()
    
    clinical2 = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
    
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
      
    print(paste0("no PC, subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("no PC, probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
    
    probes = colnames(methyDF)
    
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$plate + cF$gestage + 
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".vars.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

  }
}

```

```{bash}

cd ~/ewas/birthorder

sed -i 's/Std. Error/SE/g' *.vars.ewas.txt

### 450K birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.vars.ewas.txt
PROCESS noPC.eur.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.450K.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set1.birth_order.vars.ewas.txt
PROCESS noPC.lat.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.450K.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.vars.ewas.txt
PROCESS noPC.eur.set2.birth_order.vars.ewas.txt
PROCESS noPC.lat.set1.birth_order.vars.ewas.txt
PROCESS noPC.lat.set2.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.450K.birth_order.vars.ewas.txt

### EPIC birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set3.birth_order.vars.ewas.txt
PROCESS noPC.eur.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.EPIC.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set3.birth_order.vars.ewas.txt
PROCESS noPC.lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.EPIC.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set3.birth_order.vars.ewas.txt
PROCESS noPC.eur.set4.birth_order.vars.ewas.txt
PROCESS noPC.lat.set3.birth_order.vars.ewas.txt
PROCESS noPC.lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.EPIC.birth_order.vars.ewas.txt

### ALL birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.vars.ewas.txt
PROCESS noPC.eur.set2.birth_order.vars.ewas.txt
PROCESS noPC.eur.set3.birth_order.vars.ewas.txt
PROCESS noPC.eur.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.all.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set1.birth_order.vars.ewas.txt
PROCESS noPC.lat.set2.birth_order.vars.ewas.txt
PROCESS noPC.lat.set3.birth_order.vars.ewas.txt
PROCESS noPC.lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.all.birth_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.vars.ewas.txt
PROCESS noPC.eur.set2.birth_order.vars.ewas.txt
PROCESS noPC.eur.set3.birth_order.vars.ewas.txt
PROCESS noPC.eur.set4.birth_order.vars.ewas.txt
PROCESS noPC.lat.set1.birth_order.vars.ewas.txt
PROCESS noPC.lat.set2.birth_order.vars.ewas.txt
PROCESS noPC.lat.set3.birth_order.vars.ewas.txt
PROCESS noPC.lat.set4.birth_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.all.birth_order.vars.ewas.txt



### 450K preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.vars.ewas.txt
PROCESS noPC.eur.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.450K.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set1.preg_order.vars.ewas.txt
PROCESS noPC.lat.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.450K.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.vars.ewas.txt
PROCESS noPC.eur.set2.preg_order.vars.ewas.txt
PROCESS noPC.lat.set1.preg_order.vars.ewas.txt
PROCESS noPC.lat.set2.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.450K.preg_order.vars.ewas.txt

### EPIC preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set3.preg_order.vars.ewas.txt
PROCESS noPC.eur.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.EPIC.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set3.preg_order.vars.ewas.txt
PROCESS noPC.lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.EPIC.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set3.preg_order.vars.ewas.txt
PROCESS noPC.eur.set4.preg_order.vars.ewas.txt
PROCESS noPC.lat.set3.preg_order.vars.ewas.txt
PROCESS noPC.lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.EPIC.preg_order.vars.ewas.txt

### ALL preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.vars.ewas.txt
PROCESS noPC.eur.set2.preg_order.vars.ewas.txt
PROCESS noPC.eur.set3.preg_order.vars.ewas.txt
PROCESS noPC.eur.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.eur.meta.all.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.lat.set1.preg_order.vars.ewas.txt
PROCESS noPC.lat.set2.preg_order.vars.ewas.txt
PROCESS noPC.lat.set3.preg_order.vars.ewas.txt
PROCESS noPC.lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.lat.meta.all.preg_order.vars.ewas.txt

metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.vars.ewas.txt
PROCESS noPC.eur.set2.preg_order.vars.ewas.txt
PROCESS noPC.eur.set3.preg_order.vars.ewas.txt
PROCESS noPC.eur.set4.preg_order.vars.ewas.txt
PROCESS noPC.lat.set1.preg_order.vars.ewas.txt
PROCESS noPC.lat.set2.preg_order.vars.ewas.txt
PROCESS noPC.lat.set3.preg_order.vars.ewas.txt
PROCESS noPC.lat.set4.preg_order.vars.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.byrace.meta.all.preg_order.vars.ewas.txt


```

```{r annotation}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(qvalue)

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")
anno       <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC       <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)

for(platform in c("450K","EPIC","all")){
  for(birth in c("preg_order","birth_order")){
    for(race in c("eur","lat","byrace")){

  meta = fread(paste0("noPC.",race,".meta.",platform,".",birth,".vars.ewas.txt")) %>%
    select(-Allele1, -Allele2) %>%
    mutate(cpg = gsub("_.*$","",MarkerName),
           snp = gsub("^.*_","",MarkerName)) %>%
    arrange(by=`P-value`) %>%
    mutate(qvalue = qvalue(`P-value`)$qvalues,
           bonferroni = p.adjust(`P-value`, method = "bonferroni", n = length(`P-value`)),
           fdr = p.adjust(`P-value`, method = "fdr", n = length(`P-value`))) %>%
    filter(fdr < 0.05) %>%
    left_join(anno,by="cpg") 

  print("noPC, ",paste0(nrow(meta)," probes are significant for platform ",platform," for variable ",birth))

  if(nrow(meta)>0) write.csv(meta,paste0("noPC.",birth,".",platform,".sig.vars.meta.csv"))
  
    }}}

```

# birthOrder analysis correcting for birth related variables, by ancestry, no ancestry PCs, after talking with Stephenie London

```{r ewas of birthorder, correcting for vars}

#variables to include in the regression:
##proportional birthweight = birthweight/gestational age ?
##selection factor, caco
##maternal smoking: smoke_mo_preg
##plate, or batch
##sex
##10 cell proportion PCs/ cell proportions

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(DescTools)

# load birthorder data 

for(var in c("birth_order","preg_order")){
  
  # var = "birth_order"

  for(set in 1:2){
  
    #set=1
  
   if(set%in%c(1,2)){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       inner_join(bo, by="beadPosition") 
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==3){
    
    bo = read.csv("~/ewas/birthorder/sourceF/birthOrderRawS3.csv") %>%
       select(subjectId, birth_order, preg_order)
         
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>% 
       filter(Trisomy21==0)%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=Y_birthweight,
              gestage=Y_gestation_week)%>%
       inner_join(bo, by="subjectId")
     
     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
     
     }
    
   if(set==4){
     
      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       mutate(birth_order = as.numeric(as.character(parity)), 
              preg_order = as.numeric(as.character(mo_gravid_int))) %>%
       select(beadPosition, birth_order, preg_order)
  
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       filter(smp_type=="bg")%>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       mutate(birthWt=ch_birthwt_bc_full,
              gestage=ch_gestage_int)%>%
       inner_join(bo, by="beadPosition")

     clinical_eur = clinical0 %>% dplyr::filter(race==1)
     clinical_lat = clinical0 %>% dplyr::filter(race==3)
    
     }
  
    # load methylation data
    
    methy0 = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy0) = c()

    #methyt = methy0[1:100,1:100]
    
    # Winsorize beta values
    methy<-lapply(methy0[,-1], Winsorize, probs = c(0.005, 0.995), na.rm = FALSE) %>%
      do.call(cbind,.) %>%
      as.data.frame()
    
    methy$probeId = methy0$probeId

  ################################################################    
  # check correlation between birthWt and gestage    
  print("correlation between gestage and birth weight in Eur")  
  cor.test(clinical_eur$gestage,clinical_eur$birthWt)
  print("correlation between gestage and birth weight in Lat")  
  cor.test(clinical_lat$gestage,clinical_lat$birthWt)    
  ################################################################    
    
    # EUR  
    race="eur"
    
    clinical1 = clinical_eur %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,caco,sex,plate,birthWt,gestage,smoke_mo_preg,one_of(var)) %>%
      na.omit()
    
    clinical2 = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
  
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
    
    print(paste0("no PC, subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("no PC, probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
      
    probes = colnames(methyDF)
  
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg +
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

    ###### binary variable
    
    cF$bina = 1
    cF[which(cF[,var]==1),"bina"] = 0 
    
    reg1 <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF$bina + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".binary.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  
        
  ################################################################    
  # gestage only
    
    reg <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF[ ,var] + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".gestaOnly.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  

  reg1 <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF$bina + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".gestaOnly.binary.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  
  
  ################################################################    
  # birthweight only
  
    reg <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF[ ,var] + cF$sex+ cF$caco + cF$plate + cF$birthWt + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".birthWtOnly.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  

  reg1 <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF$bina + cF$sex + cF$caco+ cF$plate + cF$birthWt + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".birthWtOnly.binary.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  
    
  
##############################################################################################################
##############################################################################################################
    
    # LAT  
    race="lat"
    
    clinical1 = clinical_lat %>% 
      mutate(sex = as.factor(sex)) %>%
      select(beadPosition,subjectId,caco,sex,plate,birthWt,gestage,smoke_mo_preg,one_of(var)) %>%
      na.omit()
    
    clinical2 = read.csv(paste0("~/sets/set",set,"/glintControl.csv")) %>%
      select(beadPosition,rc1,rc2,rc3,rc4,rc5,rc6,rc7,rc8,rc9,rc10) %>%
      inner_join(clinical1, by="beadPosition")
    
    subs = intersect(colnames(methy)[-1],as.character(clinical2$beadPosition))
    
    methyDF = methy[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      t()%>%
      as.data.frame()
    
    clinical3 = clinical2 %>%
      column_to_rownames(var="beadPosition")
    
    cF = clinical3[subs, ]
      
    print(paste0("no PC, subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("no PC, probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
    
    probes = colnames(methyDF)
    
    reg <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF[ ,var] + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

  ###### binary variable
    
    cF$bina = 1
    cF[which(cF[,var]==1),"bina"] = 0 
    
    reg1 <- lapply(probes,function(probe){
      c(data.frame(CpG=probe),summary(
        lm(methyDF[,probe] ~ cF$bina + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
             cF$birthWt + cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
             cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
    
    reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
    
    write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".binary.ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)      
    
    
  ################################################################    
  # gestage only
    
    reg <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF[ ,var] + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".gestaOnly.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  

  reg1 <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF$bina + cF$sex+ cF$caco + cF$plate + cF$gestage + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".gestaOnly.binary.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  
  
  ################################################################    
  # birthweight only
  
    reg <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco+ cF$plate + cF$birthWt + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg_1,paste0("noPC.",race,".set",set,".",var,".birthWtOnly.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  

  reg1 <- lapply(probes,function(probe){
    c(data.frame(CpG=probe),summary(
      lm(methyDF[,probe] ~ cF$bina + cF$sex+ cF$caco + cF$plate + cF$birthWt + cF$smoke_mo_preg + 
           cF$rc1 + cF$rc2 + cF$rc3 + cF$rc4 + cF$rc5 + cF$rc6 + 
           cF$rc7 + cF$rc8 + cF$rc9 + cF$rc10))$coefficients[2,c(1,2,4)])})
  
  reg1_1 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
  
  write.table(reg1_1,paste0("noPC.",race,".set",set,".",var,".birthWtOnly.binary.ewas.txt"),
              quote=FALSE,
              sep='\t',
              row.names = FALSE, 
              col.names = TRUE)  
 
  }
}




```

#linear
noPC.eur.set1.birth_order.ewas.txt
noPC.eur.set2.birth_order.ewas.txt
noPC.lat.set1.birth_order.ewas.txt
noPC.lat.set2.birth_order.ewas.txt

noPC.eur.set1.preg_order.ewas.txt
noPC.eur.set2.preg_order.ewas.txt
noPC.lat.set1.preg_order.ewas.txt
noPC.lat.set2.preg_order.ewas.txt

#logistic
noPC.eur.set1.birth_order.binary.ewas.txt
noPC.eur.set2.birth_order.binary.ewas.txt
noPC.lat.set1.birth_order.binary.ewas.txt
noPC.lat.set2.birth_order.binary.ewas.txt

noPC.eur.set1.preg_order.binary.ewas.txt
noPC.eur.set2.preg_order.binary.ewas.txt
noPC.lat.set1.preg_order.binary.ewas.txt
noPC.lat.set2.preg_order.binary.ewas.txt

#gestage only
noPC.eur.set1.birth_order.gestaOnly.ewas.txt
noPC.eur.set2.birth_order.gestaOnly.ewas.txt
noPC.lat.set1.birth_order.gestaOnly.ewas.txt
noPC.lat.set2.birth_order.gestaOnly.ewas.txt

noPC.eur.set1.preg_order.gestaOnly.ewas.txt
noPC.eur.set2.preg_order.gestaOnly.ewas.txt
noPC.lat.set1.preg_order.gestaOnly.ewas.txt
noPC.lat.set2.preg_order.gestaOnly.ewas.txt

#birthweight only
noPC.eur.set1.birth_order.birthWtOnly.ewas.txt
noPC.eur.set2.birth_order.birthWtOnly.ewas.txt
noPC.lat.set1.birth_order.birthWtOnly.ewas.txt
noPC.lat.set2.birth_order.birthWtOnly.ewas.txt

noPC.eur.set1.preg_order.birthWtOnly.ewas.txt
noPC.eur.set2.preg_order.birthWtOnly.ewas.txt
noPC.lat.set1.preg_order.birthWtOnly.ewas.txt
noPC.lat.set2.preg_order.birthWtOnly.ewas.txt

```{bash}

cd ~/ewas/birthorder

sed -i 's/Std. Error/SE/g' *.ewas.txt

### 450K birth_order linear
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.ewas.txt
PROCESS noPC.eur.set2.birth_order.ewas.txt
PROCESS noPC.lat.set1.birth_order.ewas.txt
PROCESS noPC.lat.set2.birth_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.birth_order.lin.ewas.txt

### 450K preg_order linear
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.ewas.txt
PROCESS noPC.eur.set2.preg_order.ewas.txt
PROCESS noPC.lat.set1.preg_order.ewas.txt
PROCESS noPC.lat.set2.preg_order.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.preg_order.lin.ewas.txt

### 450K birth_order binary
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.binary.ewas.txt
PROCESS noPC.eur.set2.birth_order.binary.ewas.txt
PROCESS noPC.lat.set1.birth_order.binary.ewas.txt
PROCESS noPC.lat.set2.birth_order.binary.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.birth_order.bin.ewas.txt

### 450K preg_order binary
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.binary.ewas.txt
PROCESS noPC.eur.set2.preg_order.binary.ewas.txt
PROCESS noPC.lat.set1.preg_order.binary.ewas.txt
PROCESS noPC.lat.set2.preg_order.binary.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.preg_order.bin.ewas.txt


#gestage only, birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.birth_order.gestaOnly.ewas.txt
PROCESS noPC.eur.set2.birth_order.gestaOnly.ewas.txt
PROCESS noPC.lat.set1.birth_order.gestaOnly.ewas.txt
PROCESS noPC.lat.set2.birth_order.gestaOnly.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.birth_order.gestaOnly.bin.ewas.txt

#gestage only, preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCESS noPC.eur.set1.preg_order.gestaOnly.ewas.txt
PROCESS noPC.eur.set2.preg_order.gestaOnly.ewas.txt
PROCESS noPC.lat.set1.preg_order.gestaOnly.ewas.txt
PROCESS noPC.lat.set2.preg_order.gestaOnly.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.preg_order.gestaOnly.bin.ewas.txt

#birthweight only, birth_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCES noPC.eur.set1.birth_order.birthWtOnly.ewas.txt
PROCES noPC.eur.set2.birth_order.birthWtOnly.ewas.txt
PROCES noPC.lat.set1.birth_order.birthWtOnly.ewas.txt
PROCES noPC.lat.set2.birth_order.birthWtOnly.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.birth_order.birthWtOnly.bin.ewas.txt

#birthweight only, preg_order
metal
MARKER CpG
EFFECT Estimate
PVALUE `Pr(>|t|)`
SCHEME STDERR
STDERR SE
VERBOSE OFF
GENOMICCONTROL OFF
STDERR_PRINT_PRECISION 20
EFFECT_PRINT_PRECISION 20
PROCES noPC.eur.set1.preg_order.birthWtOnly.ewas.txt
PROCES noPC.eur.set2.preg_order.birthWtOnly.ewas.txt
PROCES noPC.lat.set1.preg_order.birthWtOnly.ewas.txt
PROCES noPC.lat.set2.preg_order.birthWtOnly.ewas.txt
ANALYZE
QUIT
mv METAANALYSIS1.TBL noPC.meta.450K.preg_order.birthWtOnly.bin.ewas.txt
```

```{r annotation}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(qvalue)

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")
anno       <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name)



for(method in c("lin","bin","birthWtOnly.bin","gestaOnly.bin")){
  for(birth in c("preg_order","birth_order")){

  meta = fread(paste0("noPC.meta.450K.",birth,".",method,".ewas.txt")) %>%
    select(-Allele1, -Allele2) %>%
    mutate(cpg = gsub("_.*$","",MarkerName),
           snp = gsub("^.*_","",MarkerName)) %>%
    arrange(by=`P-value`) %>%
    mutate(qvalue = qvalue(`P-value`)$qvalues,
           bonferroni = p.adjust(`P-value`, method = "bonferroni", n = length(`P-value`)),
           fdr = p.adjust(`P-value`, method = "fdr", n = length(`P-value`))) %>%
    filter(fdr < 0.05) %>%
    left_join(anno,by="cpg") %>%
    as.data.frame()

  paste0(nrow(meta)," probes are significant for method ",method," for variable ",birth)

  if(nrow(meta)>0) write.csv(meta,paste0("noPC.meta.450K.",birth,".",method,".sig.ewas.txt"))
  
    }}


method="bin"
birth="preg_order"

meta = fread(paste0("noPC.meta.450K.",birth,".",method,".ewas.txt")) %>%
  select(-Allele1, -Allele2) %>%
  mutate(cpg = gsub("_.*$","",MarkerName),
         snp = gsub("^.*_","",MarkerName)) %>%
  arrange(by=`P-value`) %>%
  left_join(anno,by="cpg") %>%
  as.data.frame()
metaTop = meta[1:20,]
write.csv(metaTop,"topPreg_orderCpGsBinary.csv")

birth="birth_order"

meta = fread(paste0("noPC.meta.450K.",birth,".",method,".ewas.txt")) %>%
  select(-Allele1, -Allele2) %>%
  mutate(cpg = gsub("_.*$","",MarkerName),
         snp = gsub("^.*_","",MarkerName)) %>%
  arrange(by=`P-value`) %>%
  left_join(anno,by="cpg") %>%
  as.data.frame()
metaTop = meta[1:20,]
write.csv(metaTop,"topBirth_orderCpGsBinary.csv")

```

[1] "0 probes are significant for method lin for variable preg_order"
[1] "0 probes are significant for method lin for variable birth_order"
[1] "0 probes are significant for method bin for variable preg_order"
[1] "0 probes are significant for method bin for variable birth_order"
[1] "0 probes are significant for method birthWtOnly.bin for variable preg_order"
[1] "0 probes are significant for method birthWtOnly.bin for variable birth_order"
[1] "0 probes are significant for method gestaOnly.bin for variable preg_order"
[1] "0 probes are significant for method gestaOnly.bin for variable birth_order"

```{r if changing covariates changes the rank}

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)

# birth_order

bthVar = fread("noPC.meta.450K.birth_order.bin.ewas.txt") %>%
  arrange(`P-value`)%>%
  mutate(rankBoth = rank(`P-value`),
         EffectBoth=Effect) %>%
  select(MarkerName,rankBoth,EffectBoth)

bthGesta = fread("noPC.meta.450K.birth_order.gestaOnly.bin.ewas.txt") %>%
  mutate(rankGesta = rank(`P-value`),
         EffectGesta=Effect) %>%
  select(MarkerName,rankGesta,EffectGesta)

bthBirW = fread("noPC.meta.450K.birth_order.birthWtOnly.bin.ewas.txt") %>%
  mutate(rankBirW = rank(`P-value`),
         EffectBirthW=Effect) %>%
  select(MarkerName,rankBirW,EffectBirthW)

bthVar1k = bthVar[1:1000,]

bthVar %>%
  inner_join(bthGesta,by="MarkerName")%>%
  ggplot(aes(x=EffectBoth,y=EffectGesta))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic() +
  ggtitle("Effects of CpGs in BO regression")+
  xlab("CpG effect in models including both GestA and BirthW") +
  ylab("CpG effect in models including only GestA") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("effectCompBothGesta.png",width = 8,height = 6)


bthVar %>%
  inner_join(bthBirW,by="MarkerName")%>%
  ggplot(aes(x=EffectBoth,y=EffectBirthW))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic() +
  ggtitle("Effects of CpGs in BO regression")+
  xlab("CpG effect in models including both GestA and BirthW") +
  ylab("CpG effect in models including only Birthweight") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("effectCompBothBirthW.png",width = 8,height = 6)

```

# birthOrder analysis correcting for birth related variables, by ancestry, no ancestry PCs, controlling for cell proportions, testing preg-weight variables

```{r ewas of birthorder, correcting for vars}

#variables to include in the regression:
##proportional birthweight = birthweight/gestational age ?
##selection factor, caco
##maternal smoking: smoke_mo_preg
##plate, or batch
##sex
##10 cell proportion PCs/ cell proportions

setwd("~/ewas/birthorder")
library(tidyverse)
library(data.table)
library(DescTools)

# load birthorder data 

#for(var in c("birth_order","preg_order")){
  
  var = "birth_order"

  for(set in c(1,2)){
  

      setN = paste0("set",set)
        
      bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
       filter(set == setN)%>%
       select(beadPosition, birth_order, preg_order)
    
     clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
       mutate(plate = as.factor(gsub("_.*$","",beadPosition))) %>%
       inner_join(bo, by="beadPosition") 
    
    # load methylation data
    
    methy0 = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds"))
    rownames(methy0) = c()

    # Winsorize beta values
    methy<-lapply(methy0[,-1], Winsorize, probs = c(0.005, 0.995), na.rm = FALSE) %>%
      do.call(cbind,.) %>%
      as.data.frame()
    
    methy$probeId = methy0$probeId

    for(race in c("eur","lat")){
        
      if(race=="eur") clinical_sub = clinical0 %>% dplyr::filter(race==1)
      if(race=="lat") clinical_sub = clinical0 %>% dplyr::filter(race==3)
      
      clinical1 = clinical_sub %>% 
        mutate(sex = as.factor(sex)) %>%
        select(beadPosition,subjectId,caco,sex,plate,birthWt,
               gestage,smoke_mo_preg,mo_age,one_of(var))
      
      clinical2 = read.csv(paste0("~/sets/set",set,"/idol_deconvolution.csv")) %>%
        select(beadPosition,CD8T,CD4T,NK,Bcell,Mono,Gran,nRBC) %>%
        inner_join(clinical1, by="beadPosition")
    
      setN=set
      clinical3 = read.csv("~/ewas/birthorder/sourceF/set124_weightgain.csv") %>%
        filter(set==paste0("set",setN)) %>%
        select(beadPosition,mo_weightgain,mo_weight)%>%
        inner_join(clinical2, by="beadPosition")

      subs = intersect(colnames(methy)[-1],as.character(clinical3$beadPosition))
      
      methyDF = methy[,c("probeId",subs)] %>%
        column_to_rownames(var="probeId")%>%
        t()%>%
        as.data.frame()
    
      clinical3 = clinical3 %>%
        column_to_rownames(var="beadPosition") %>%
        na.omit()
      
      cF = clinical3[subs, ]
      
      print(paste0("subject number in set ",set," of race ",race, " is ", nrow(cF)))  
      print(paste0("probe number in set ",set," of race ",race, " is ", ncol(methyDF)))  
        
      probes = colnames(methyDF)
    
      # model 1
      reg <- lapply(probes,function(probe){
        c(data.frame(CpG=probe),summary(
          lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco + 
               cF$plate + cF$gestage + cF$mo_age + cF$smoke_mo_preg + 
               cF$birthWt + cF$CD8T + cF$CD4T + cF$NK + cF$Bcell + 
               cF$Mono + cF$Gran + cF$nRBC))$coefficients[2,c(1,2,4)])})
      reg_1 <- bind_rows(reg) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
      write.table(reg_1,paste0("sensitivitytest.nogain.noweight.",race,".set",set,".",var,".ewas.txt"),
                  quote=FALSE,
                  sep='\t',
                  row.names = FALSE, 
                  col.names = TRUE)  
      
        # model 2
      reg1 <- lapply(probes,function(probe){
        c(data.frame(CpG=probe),summary(
          lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco + cF$mo_weight + 
               cF$plate + cF$gestage + cF$mo_age + cF$smoke_mo_preg + 
               cF$birthWt + cF$CD8T + cF$CD4T + cF$NK + cF$Bcell + 
               cF$Mono + cF$Gran + cF$nRBC))$coefficients[2,c(1,2,4)])})
      reg_2 <- bind_rows(reg1) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
      write.table(reg_2,paste0("sensitivitytest.nogain.weight.",race,".set",set,".",var,".ewas.txt"),
                  quote=FALSE,
                  sep='\t',
                  row.names = FALSE, 
                  col.names = TRUE)  
      
        # model 3
      reg2 <- lapply(probes,function(probe){
        c(data.frame(CpG=probe),summary(
          lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco + cF$mo_weightgain + 
               cF$plate + cF$gestage + cF$mo_age + cF$smoke_mo_preg + 
               cF$birthWt + cF$CD8T + cF$CD4T + cF$NK + cF$Bcell + 
               cF$Mono + cF$Gran + cF$nRBC))$coefficients[2,c(1,2,4)])})
      reg_3 <- bind_rows(reg2) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
      write.table(reg_3,paste0("sensitivitytest.gain.noweight.",race,".set",set,".",var,".ewas.txt"),
                  quote=FALSE,
                  sep='\t',
                  row.names = FALSE, 
                  col.names = TRUE)  
      
        # model 4
      reg3 <- lapply(probes,function(probe){
        c(data.frame(CpG=probe),summary(
          lm(methyDF[,probe] ~ cF[ ,var] + cF$sex + cF$caco + cF$mo_weight +
               cF$mo_weightgain + 
               cF$plate + cF$gestage + cF$mo_age + cF$smoke_mo_preg + 
               cF$birthWt + cF$CD8T + cF$CD4T + cF$NK + cF$Bcell + 
               cF$Mono + cF$Gran + cF$nRBC))$coefficients[2,c(1,2,4)])})
      reg_4 <- bind_rows(reg3) %>% as.data.frame() %>% arrange(by=`Pr(>|t|)`)
      write.table(reg_4,paste0("sensitivitytest.gain.weight.",race,".set",set,".",var,".ewas.txt"),
                  quote=FALSE,
                  sep='\t',
                  row.names = FALSE, 
                  col.names = TRUE)  

    }  
  }
#}




```

[1] "subject number in set 1 of race eur is 199"
[1] "probe number in set 1 of race eur is 483287"
[1] "subject number in set 1 of race lat is 170"
[1] "probe number in set 1 of race lat is 483287"
[1] "subject number in set 2 of race eur is 148"
[1] "probe number in set 2 of race eur is 484552"
[1] "subject number in set 2 of race lat is 219"
[1] "probe number in set 2 of race lat is 484552"

```{r comparing results}

library(tidyverse)
library(data.table)
library(cowplot)

setwd("~/ewas/birthorder")

library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
data("IlluminaHumanMethylation450kanno.ilmn12.hg19")

for(set in c(1,2)){
  for(race in c("eur","lat")){
    
  #no weight no gain
  nwng = fread(paste0("sensitivitytest.nogain.noweight.",race,".set",set,".birth_order.ewas.txt"))
  nwng500 = nwng[1:500]
  
  #weight, no gain
  wng = fread(paste0("sensitivitytest.nogain.weight.",race,".set",set,".birth_order.ewas.txt"))
  wng500 = wng[1:500]

  #no weight, gain
  nwg = fread(paste0("sensitivitytest.gain.noweight.",race,".set",set,".birth_order.ewas.txt"))
  nwg500 = nwg[1:500]
  
  #weight, and gain
  wg = fread(paste0("sensitivitytest.gain.weight.",race,".set",set,".birth_order.ewas.txt"))
  wg500 = wg[1:500]
  
  ## Annotate top 10 CpGs and genes in each dataset
  anno <- IlluminaHumanMethylation450kanno.ilmn12.hg19 %>%
    getAnnotation %>%
    as.data.frame %>%
    rownames_to_column(var="CpG") %>%
    select(CpG,UCSC_RefGene_Name)
  
  nwng10 = nwng500[1:10] %>% select(CpG,`Pr(>|t|)`) %>% left_join(anno,by=c("CpG")) %>% mutate(model="nwng")
  wng10 = wng500[1:10] %>% select(CpG,`Pr(>|t|)`) %>% left_join(anno,by=c("CpG")) %>% mutate(model="wng")
  nwg10 = nwg500[1:10] %>% select(CpG,`Pr(>|t|)`) %>% left_join(anno,by=c("CpG")) %>% mutate(model="nwg")
  wg10 = wg500[1:10] %>% select(CpG,`Pr(>|t|)`) %>% left_join(anno,by=c("CpG")) %>% mutate(model="wg")
  
  topComp = cbind(nwng10,wng10,nwg10,wg10) %>% as.data.frame()
  
  write.csv(topComp,paste0("comparison.of.top.hits.",race,".set",set,".csv"))
  
  ## Correlation between beta value and P values of top 500 hits
  
  com="beta values"
  model0 = "no weight no gain"
  modelX = "weight no gain"
  
  iris1= nwng500 %>%
    inner_join(wng500,by="CpG") %>%
    ggplot(aes(x=Estimate.x,y=Estimate.y))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))

  com="p values"
  model0 = "no weight no gain"
  modelX = "weight no gain"
  
  iris2= nwng500 %>%
    inner_join(wng500,by="CpG") %>%
    ggplot(aes(x=`Pr(>|t|).x`,y=`Pr(>|t|).y`))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))


  com="beta values"
  model0 = "no weight no gain"
  modelX = "gain but no weight  "
  
  iris3= nwng500 %>%
    inner_join(nwg500,by="CpG") %>%
    ggplot(aes(x=Estimate.x,y=Estimate.y))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))

  com="p values"
  model0 = "no weight no gain"
  modelX = "gain but no weight  "
  
  iris4= nwng500 %>%
    inner_join(nwg500,by="CpG") %>%
    ggplot(aes(x=`Pr(>|t|).x`,y=`Pr(>|t|).y`))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))


  com="beta values"
  model0 = "no weight no gain"
  modelX = "weight and gain"
  
  iris5= nwng500 %>%
    inner_join(wg500,by="CpG") %>%
    ggplot(aes(x=Estimate.x,y=Estimate.y))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))

  com="p values"
  model0 = "no weight no gain"
  modelX = "weight and gain"
  
  iris6= nwng500 %>%
    inner_join(wg500,by="CpG") %>%
    ggplot(aes(x=`Pr(>|t|).x`,y=`Pr(>|t|).y`))+
    geom_point()+
    geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
    ggtitle(paste0("comparison between ",model0, " and ", modelX,"\n",com, " ",race," in set ",set))+
    xlab(paste0("beta values ",model0)) + ylab(paste0("beta values ",modelX))+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))

plot_grid(iris1, iris2, iris3, iris4, iris5, iris6, labels = "AUTO",nrow=3)
ggsave(paste0("sensitivityAnalysis.",race,".set",set,".png"),width = 13,height = 15)
  

  }
}  


```

# birthorder, testing robust linear regression codes with annotation

```{r}
#packages <- c("tidyverse", "data.table", "DescTools", "MASS", "lmtest", "sandwich")
#install.packages(setdiff(packages, rownames(installed.packages())))  

setwd("~/ewas/birthorder") # your working directory
library(tidyverse)
library(data.table)
library(DescTools)
library(MASS)
library(lmtest)
library(sandwich)
library(tidyverse)
library(parallel) 

# Each race should have its own model and csv result file. Do not include any race with fewer than 15 subjects.
# Each array should have its own model and csv result file. 
# If you have miscarriage/abortion data, you should run two models for each situation (including and excluding miscarriage/abortion) 
# Run the following for each model, replacing appropriate file names
# Winsorize top and bottom beta values

RLM.Robust <- function(meth, cov){

    # meth = methyl["cg25813447",]
    # cov = covs
      
    dat <- data.frame(y = meth, cov)
    ff  <- formula(paste("y ~ ", paste0(colnames(cov), collapse="+")))
    mod <- try(rlm(ff, data = dat, maxit = 200))
    #cf = try(coeftest(mod, vcov=vcovHC(mod, type="HC0")))
    cf = try(coeftest(mod))

  	if (exists(class(cf)) == FALSE) {
  		out = rep(NA,3)
  	} else if (exists(class(cf)) == TRUE) {
  		if( class(cf)[1] == "try-error") {
  		cat("error throw by CpG", set, ";")
  		out = rep(NA,3)
  		print(paste0(cov, " - chol2inv could not be computed"))
  		} 
  		else {
  			coef = cf[2,"Estimate"]
  			se = cf[2,"Std. Error"]
  			p.value = cf[2,"Pr(>|z|)"]
  			out = c(coef, se, p.value) 
  		}
    }
  names(out) = c("Beta","SE", "P"); return(out) 
  }

winsorize <- function(methylation,pct=winsorize.pct) {
  quantiles <- matrixStats::rowQuantiles(methylation, probs=c(pct,1-pct), na.rm=T)
  low <- quantiles[,1]
  upper <- quantiles[,2]

  outliers.lower <- rowSums(methylation < low, na.rm=T)
  outliers.upper <- rowSums(methylation > upper, na.rm=T)
  
  idx <- which(methylation < low, arr.ind=T)
  methylation[idx] <- low[idx[,1]]
  
  idx <- which(methylation > upper, arr.ind=T)
  methylation[idx] <- upper[idx[,1]]

  n <- rowSums(!is.na(methylation))
  log <- data.frame(outliers.lower, outliers.upper, n)
  return(list(methylation=methylation, log=log))
}

for(set in c(1,2)){
  
    setN = paste0("set",set)
      
    bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
     filter(set == setN)%>%
     dplyr::select(beadPosition, birth_order, preg_order) %>%
     dplyr::mutate(boCE=birth_order,
                   boC=ifelse(birth_order==1,0,1),
                   boCI=preg_order,
                   boI=ifelse(preg_order==1,0,1))
  
   clinical0 = read.csv(paste0("~/sets/set",set,"/clinical_variables.csv")) %>%
     inner_join(bo, by="beadPosition") 
  
  # load methylation data
    #methy0 = read_rds(paste0("~/sets/set",set,"/beta_noob_imputed_set",set,".rds")) %>%
    methy0 = fread(paste0("~/sets/set",set,"/set",set,"_Sesame_beta.csv")) %>%
          as.data.frame()
    methy0 = methy0[,-1]
    
    rownames(methy0) = c()

  for(race in c("eur","lat")){
    
    #race="eur"
      
    if(race=="eur") clinical_sub = clinical0 %>% dplyr::filter(race==1)
    if(race=="lat") clinical_sub = clinical0 %>% dplyr::filter(race==3)
    
    clinical1 = clinical_sub %>% 
      dplyr::mutate(sex = as.factor(sex)) %>%
      dplyr::select(beadPosition,subjectId,caco,sex,birthWt,
             gestage,smoke_mo_preg,mo_age,boCE,boC,boCI,boI)%>%
             mutate(plate = as.factor(as.character(gsub("_.*$","",beadPosition))))

    
    clinical2 = read.csv(paste0("~/sets/set",set,"/idol_deconvolution.csv")) %>%
      dplyr::select(beadPosition,CD8T,CD4T,NK,Bcell,Mono,Gran,nRBC) %>%
      dplyr::inner_join(clinical1, by="beadPosition")
  
    setN=set
    
    clinical3 = read.csv("~/ewas/birthorder/sourceF/set124_weightgain.csv") %>%
      dplyr::filter(set==paste0("set",setN)) %>%
      dplyr::select(beadPosition,mo_weightgain,mo_weight)%>%
      dplyr::inner_join(clinical2, by="beadPosition")

    subs = intersect(colnames(methy0),as.character(clinical3$beadPosition))

    clinical4 = clinical3 %>%
      column_to_rownames(var="beadPosition") 
    
    cF = clinical4[subs, ] %>% dplyr::select(-subjectId)

    methyDF = methy0[,c("probeId",subs)] %>%
      column_to_rownames(var="probeId")%>%
      as.matrix()

    replace.outliers <- winsorize(methyDF, 0.005)
    methy1 <- replace.outliers$methylation

    print(paste0("subject number in set ",set," of race ",race, " is ", nrow(cF)))  
    print(paste0("probe number in set ",set," of race ",race, " is ", nrow(methy1)))  
    dim(cF)
    dim(methy1)

    ### To run model 1A
    var= "boCE"

    covs = cF[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran")] 
    reg_1 <- t(apply(methy1, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P) 

    write.table(reg_1,paste0("bo.1A.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  
    
    ### To run model 1B
    var= "boC"

    covs = cF[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran")] 
    reg_1 <- t(apply(methy1, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P)

    write.table(reg_1,paste0("bo.1B.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  


    ### To run model 2A
    var= "boCI"

    covs = cF[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran")] 
    reg_1 <- t(apply(methy1, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P)
        
    write.table(reg_1,paste0("bo.2A.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

    ### To run model 2B
    var= "boI"

    covs = cF[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran")] 
    reg_1 <- t(apply(methy1, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P)
    
    write.table(reg_1,paste0("bo.2B.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

   ### To run model S3A
    
    cf1 = cF %>% na.omit()
    cf2 = droplevels(cf1)
    methy2 = methy1[,rownames(cf1)]

    var= "boCE"
    
    covs = cf2[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran","mo_weight")] 
    reg_1 <- t(apply(methy2, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P)

    write.table(reg_1,paste0("bo.S3A.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

   ### To run model S3B
    var= "boCE"
    
    covs = cf2[,c(var,"sex","caco","plate","gestage","mo_age","smoke_mo_preg","birthWt","nRBC","CD8T","CD4T","NK","Bcell","Mono","Gran","mo_weightgain")]
    reg_1 <- t(apply(methy2, 1, RLM.Robust, covs))  %>% 
      as.data.frame %>%
      arrange(by=P)
    
    write.table(reg_1,paste0("bo.S3B.",race,".set",set,".",var,".ewas.txt"),
                quote=FALSE,
                sep='\t',
                row.names = FALSE, 
                col.names = TRUE)  

  }  
}



```

# Meta-analysis of multiple cohorts

```{r drawing manhatton plot}

setwd("/ccls/home/sli/ewas/birthorder/meta")
library(tidyverse)
library(data.table)
library(CMplot)
library(QCEWAS)
library(qqman)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  dplyr::select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name,
                Probe_maf,UCSC_RefGene_Group,CpG_maf)

#Loadback and annotate
files = c("meta.1A.all.all.TBL",
          "meta.1B.all.all.TBL",
          "meta.2A.all.all.TBL",
          "meta.2B.all.all.TBL",
          "meta.S3A.all.all.TBL",
          "meta.S3B.all.all.TBL")

for(i in files){
  
  #i = "meta.1A.all.all.TBL"
  
  ewas_DS_ALL <- fread(i)%>%
    dplyr::rename("cpg"="MarkerName",
                  "P" = "P-value") %>%
    dplyr::filter(grepl('cg', cpg)) %>%
    filter(!is.na(P))%>% 
    left_join(annoEPIC,by="cpg") %>%
    arrange(by=P) %>% 
    dplyr::filter(chr!="chrX"&chr!="chrY") %>%  
    dplyr::filter(Probe_maf<0.05|is.na(Probe_maf)) %>% 
    dplyr::filter(CpG_maf<0.05|is.na(CpG_maf)) %>%
    mutate(adjp = p.adjust(`P`, method="fdr"),
           adjp_bon = p.adjust(`P`, method="bonferroni"))  
  
  filename = paste0(gsub("meta.|.TBL","",i),".anno.txt")
  write.table(ewas_DS_ALL,filename,sep='\t',quote=FALSE,row.names = FALSE)
      
  sig_thresghold = ewas_DS_ALL %>%
    filter(adjp<0.05)%>%
    arrange(desc(adjp))
  nrow(sig_thresghold)
  sig_threshold = sig_thresghold[1,"P"]
  
  outcomes_model_CM_plot <- ewas_DS_ALL%>%
    dplyr::mutate(p_val_cmplot = -log10(`P`))%>%
    dplyr::mutate(p_val_cmplot= ifelse(Effect<0,-p_val_cmplot, p_val_cmplot )) %>%
    dplyr::mutate(cpg_name=cpg) %>%
    dplyr::select(cpg_name,chr,pos, p_val_cmplot) %>%
    na.omit()
  outcomes_model_CM_plot$chr <- gsub("chr","",outcomes_model_CM_plot$chr)
  
  if(is.na(sig_threshold)){
  CMplot(outcomes_model_CM_plot,
         plot.type="m",
         band=0, 
         cex=c(0.5,0.5,0.5),
         amplify=FALSE,
         LOG10=FALSE,  
         signal.col=NULL,
         chr.den.col=NULL,
         file="jpg",
         memo=i,
         main=i,
         dpi=500,
         file.output=TRUE,
         verbose=TRUE)
  }else{
  CMplot(outcomes_model_CM_plot,
         plot.type="m",
         band=0, 
         cex=c(0.5,0.5,0.5),
         amplify=FALSE,
         LOG10=FALSE,  
         signal.col=NULL,
         chr.den.col=NULL,
         threshold=c(log10(sig_threshold[[1]]), -log10(sig_threshold[[1]])),
         threshold.lty=2, 
         threshold.lwd=1, 
         threshold.col="black", 
         file="jpg",
         memo=i,
         main=i,
         dpi=500,
         file.output=TRUE,
         verbose=TRUE)    
  }
  
  outcomes_model_qq <- ewas_DS_ALL %>% 
    dplyr::mutate(ID=cpg)
  lambda <- P_lambda(outcomes_model_qq$P)
  
  tiff(paste0(i,".lambda.tiff"), width = 10, height = 7, units = 'in', res = 300)
  qq(outcomes_model_qq$P, main=paste0("lambda value= ",round(lambda,3)))
  dev.off()
  
}


# specifically for model 1A
i = "meta.1A.all.all.TBL"
filename = paste0(gsub("meta.|.TBL","",i),".anno.txt")

sig_thresghold = fread(filename) %>%
  as.data.frame() %>%
  filter(adjp<0.05)%>%
  arrange(desc(adjp))
nrow(sig_thresghold)
sig_threshold = sig_thresghold[1,"P"]

annotation = ewas_DS_ALL %>%
  filter(UCSC_RefGene_Name!="") %>%
  filter( (P<=2.03E-08) | cpg %in% c("cg01176516") ) %>%
  #filter(P<=2.03E-08 | (chr=="chr7")&(pos==56515510 | pos==63506148 )) %>%
  mutate(gene.highlight = sub(";.*$","",UCSC_RefGene_Name))
cpg.highlight = annotation$cpg
genes.highlight = annotation$gene.highlight

outcomes_model_CM_plot <- ewas_DS_ALL%>%
  dplyr::mutate(p_val_cmplot = -log10(P))%>%
  dplyr::mutate(p_val_cmplot= ifelse(Effect<0,-p_val_cmplot, p_val_cmplot )) %>%
  dplyr::select(cpg, chr,pos, p_val_cmplot)
outcomes_model_CM_plot$chr <- gsub("chr","",outcomes_model_CM_plot$chr)

CMplot(outcomes_model_CM_plot,
       plot.type="m",
       band=0, 
       cex=c(0.5,0.5,0.5),
       amplify=FALSE,
       LOG10=FALSE,  
       signal.col=NULL,
       chr.den.col=NULL,
       threshold=c(log10(sig_threshold[[1]]), -log10(sig_threshold[[1]])),
       threshold.lty=2, 
       threshold.lwd=1, 
       threshold.col="black", 
       highlight=cpg.highlight,
       highlight.text=genes.highlight, 
       file="jpg",
       memo=paste0("anno.",i),
       dpi=500,
       file.output=TRUE,
       verbose=TRUE)    


```

#top CpGs and pathway analysis

```{r top CpGs and pathway analysis}

#all, all
### top 10 CpGs identified by the analysis
setwd("/ccls/home/sli/ewas/birthorder/meta")

library(tidyverse)
library(data.table)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)

data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") 

platform="all"
race="all"

for(modelN in c("1A","1B","2A","2B","S3A","S3B")){
  
  #modelN="1A"
  
  model <- fread(paste0(modelN,".",platform,".",race,".anno.txt")) %>%
    as.data.frame() 
  
  sigCpGs <- model %>%
    filter(adjp<0.05) %>%
    write_tsv(paste0("sig.cpgs.",modelN,".",platform,".",race,".txt"))
  
  sigCpGsVis <- model %>%
    filter(adjp<0.05) %>%
    filter(UCSC_RefGene_Name!="") %>%
    mutate(geneName = gsub(";.*","",UCSC_RefGene_Name))
  data.frame(sigCpGsVis$geneName)
  table(sigCpGsVis$geneName) %>% max()
  
  topsCpG <- head(model,10) 
  a = topsCpG %>% 
    dplyr::select(cpg,Direction,adjp,Relation_to_Island,UCSC_RefGene_Name)
  
  #print(modelN)
  #print(a)
  
  meta = model %>%
    arrange(adjp) %>%
    dplyr::filter(adjp<0.05) 
  dim(meta)
  
  ### pathway analysis for adj P < 0.05
  meta = model %>%
    arrange(adjp) %>%
    dplyr::filter(adjp<0.05) %>%
    dplyr::select(cpg,P,adjp) 
  write_rds(meta,paste0(modelN,".meta.rds"))

  meta = model %>%
    arrange(adjp) %>%
    dplyr::filter(adjp<0.05,HetPVal>0.05) %>%
    dplyr::select(cpg,P,adjp) 
  dim(meta)

  write_rds(meta,paste0(modelN,".cons.meta.rds"))  
}

# local computer
library(methylGSA)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(tidyverse)
setwd("/Users/bo/Downloads/before")
#setwd("/ccls/home/sli/ewas/birthorder/meta")
platform="all"
race="all"
  
for(modelN in c("1A","1B","2A","2B","S3A","S3B")){
    
  #modelN="1A"

  meta = read_rds(paste0(modelN,".meta.rds"))

  cpg.pval = as.numeric(meta$P)
  names(cpg.pval) = meta$cpg 
  
  res1 = methylglm(cpg.pval = cpg.pval, GS.type = "KEGG", array.type = "EPIC")
  res2 = methylglm(cpg.pval = cpg.pval, GS.type = "GO", array.type = "EPIC")
  
  write.table(res1,paste0("../after/pathway.kegg.",modelN,".",platform,".",race,".txt"),
              sep = "\t", row.names = FALSE, quote = FALSE)
  write.table(res2,paste0("../after/pathway.go.",modelN,".",platform,".",race,".txt"),
              sep = "\t", row.names = FALSE, quote = FALSE)

}

## consistent CpGs only
for(modelN in c("1A","1B","2A","2B","S3A","S3B")){
    
  #modelN="1A"

  meta = read_rds(paste0(modelN,".cons.meta.rds"))
  
  cpg.pval = as.numeric(meta$P)
  names(cpg.pval) = meta$cpg 
  
  res1 = methylglm(cpg.pval = cpg.pval, GS.type = "KEGG", array.type = "EPIC")
  res2 = methylglm(cpg.pval = cpg.pval, GS.type = "GO", array.type = "EPIC")
  
  write.table(res1,paste0("../after/consistent.pathway.kegg.",modelN,".",platform,".",race,".txt"),
              sep = "\t", row.names = FALSE, quote = FALSE)
  write.table(res2,paste0("../after/consistent.pathway.go.",modelN,".",platform,".",race,".txt"),
              sep = "\t", row.names = FALSE, quote = FALSE)

}
```

#DMR analysis from the analysis

```{r DMRs}

setwd("/ccls/home/sli/ewas/birthorder/meta")
library(tidyverse)
library(data.table)
library(ENmix)
library(GenomicRanges)
library(AnnotationHub)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(annotatr)
library(methyAnalysis)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)

platform="all"
race="all"

data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  mutate(start=pos,
         end=pos,
         probe=Name) %>%
  dplyr::select(chr,start,end,probe) 

for(modelN in c("1A","1B","2A","2B","S3A","S3B")){

 # modelN = "1A"  
 raw = fread(paste0(modelN,".",platform,".",race,".anno.txt")) %>%
    as.data.frame() %>%
    dplyr::mutate(start=pos, end=pos) 
  
 raw1 = raw %>%
    dplyr::rename("probe"="cpg",
                  "p"="P") %>%
    dplyr::select(probe,p,chr,start,end) %>%
    mutate(chr=as.numeric(sub("chr","",chr)))%>%
    arrange(chr,start)%>%
    filter(p > 0 & p <1) %>%
    na.omit() 
  
  ipdmr(raw1,
        include.all.sig.sites=FALSE,
        region_plot=FALSE,
        mht_plot=FALSE)
  
  # if (file.exists("region_plot.pdf")) {
  #  file.rename("region_plot.pdf", paste0(modelN,".",platform,".",race,"_region_plot.pdf"))
  # } 
  
  if(file.exists("resu_ipdmr.csv")){
  
    dmr = read.csv("resu_ipdmr.csv") %>%
      filter(nprobe>1)
    
    comb_p_GR <- makeGRangesFromDataFrame(dmr, keep.extra.columns=T,
                             ignore.strand=T,
                             seqinfo=NULL,
                             seqnames.field=c("chr"),
                             start.field="start",
                             end.field="end",
                             strand.field="strand",
                             starts.in.df.are.0based=FALSE)
    
    if (require(TxDb.Hsapiens.UCSC.hg19.knownGene)) {
    try(comb_p_GR_DF <-annotateDMRInfo(comb_p_GR, 'TxDb.Hsapiens.UCSC.hg19.knownGene'),
        silent=TRUE)
    }
    
    comb_p_GR_DF <- data.frame(comb_p_GR_DF$sigDMRInfo) %>%
      mutate(dmr_id = paste0(seqnames,"_",start,"_",end))%>%
      dplyr::select(-strand)
    
    dmrs_1 = comb_p_GR_DF %>%  
      dplyr::select(seqnames,start,end,dmr_id) %>%
      as.data.frame()
  
    # betasAll = raw %>%
    #   filter(!is.na(P))%>%
    #   #mutate(CpG = V1) %>%
    #   dplyr::select(cpg,Effect) %>%
    #   left_join(annoEPIC,by=c("cpg"="probe")) %>%
    #   arrange(chr,start)%>%
    #   na.omit()
          
    betasAll = raw %>%
      filter(!is.na(P))%>%
      dplyr::select(cpg,Effect,chr,start,end) %>%
      arrange(gsub("chr","",chr),start)%>%
      na.omit()
    
    DMR_GR <- makeGRangesFromDataFrame(dmrs_1, keep.extra.columns=T,
                           ignore.strand=T,
                           seqinfo=NULL,
                           seqnames.field=c("seqnames"),
                           start.field="start",
                           end.field=c("end"),
                           strand.field="strand",
                           starts.in.df.are.0based=FALSE)
    Res_GR <- makeGRangesFromDataFrame(betasAll, keep.extra.columns=T,
                             ignore.strand=T,
                             seqinfo=NULL,
                             seqnames.field=c("chr"),
                             start.field="start",
                             end.field=c("end"),
                             strand.field="strand",
                             starts.in.df.are.0based=FALSE)
    overlap_DMR_Res <- GenomicRanges::findOverlaps(Res_GR,DMR_GR, ignore.strand=T)
    
    a=dmrs_1[subjectHits(overlap_DMR_Res),] %>%
      dplyr::select(dmr_id)
    b=betasAll[queryHits(overlap_DMR_Res),] 
    com = cbind(b,a) %>% 
      as.data.frame() %>%
      arrange(chr,start) %>%
      group_by(dmr_id) %>%
      summarise(mean_beta = mean(Effect))
    
    dmr_w_beta = comb_p_GR_DF %>%
      left_join(com,by="dmr_id") %>%
      dplyr::select(-dmr_id)
    
    write.csv(dmr_w_beta,paste0(modelN,".",platform,".",race,".annotated_dmr.csv"))
    
    unlink("resu_ipdmr.csv") # delete this file
  
  }
}

```

```{r Epigenetic score for birthweight (LASSO regression)}

setwd("/ccls/home/sli/ewas/birthorder/meta/score")
library(tidyverse)
library(data.table)

library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  dplyr::select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name,
                Probe_maf,UCSC_RefGene_Group,CpG_maf)

# Top CpGs for epigenetic part of birth order results

i = "meta.1A.all.all.no.usc.TBL"

ewas_DS_ALL <- fread(paste0("../",i))%>%
  dplyr::rename("cpg"="MarkerName",
                "P" = "P-value") %>%
  dplyr::filter(grepl('cg', cpg)) %>%
  filter(!is.na(P))%>% 
  left_join(annoEPIC,by="cpg") %>%
  arrange(by=P) %>% 
  dplyr::filter(chr!="chrX"&chr!="chrY") %>%  
  dplyr::filter(Probe_maf<0.05|is.na(Probe_maf)) %>% 
  dplyr::filter(CpG_maf<0.05|is.na(CpG_maf)) %>%
  mutate(adjp = p.adjust(`P`, method="fdr"),
         adjp_bon = p.adjust(`P`, method="bonferroni")) %>%
  dplyr::select(cpg,Effect)

ewas_DS_ALL_100 = ewas_DS_ALL[1:100,] 

write_rds(ewas_DS_ALL_100,"top.100.bo.score.rds")

# read set 2 methylation data and compare case control diffrence
ewas_DS_ALL_100 = read_rds("top.100.bo.score.rds") %>%
  mutate(Effect_abs = abs(Effect))

bo = read.csv("~/ewas/birthorder/sourceF/bthorder_meth_ph123_neo.csv") %>%
 filter(set == "set2")%>%
 dplyr::select(beadPosition, birth_order)

s2_clinical = read.csv("/ccls/home/sli/sets/set2/clinical_variables.csv") %>%
  filter(!is.na(caco))%>%
  #filter(leuk_dx_type=="B"| CaCo==0) %>%
  left_join(bo, by="beadPosition")

s2 = read_rds("/ccls/home/sli/sets/set2/set2_Sesame_beta_imputed.rds") %>%
  filter(probeId %in% ewas_DS_ALL_100$cpg) %>%
  column_to_rownames(var="probeId") %>%
  t %>%
  as.matrix()
  
coef = ewas_DS_ALL_100 %>%
  column_to_rownames(var="cpg")  

coef_m = coef[colnames(s2),,drop=FALSE] %>% as.matrix()
colnames(coef_m) = c("bo_score","bo_score_abs")

score = s2 %*% coef_m[,1] %>% # 443X78 * 78X1 
  as.data.frame() %>%
  dplyr::rename("bo_score"="V1") %>%
  rownames_to_column(var="beadPosition") %>%
  inner_join(s2_clinical,by="beadPosition")

score_abs = s2 %*% coef_m[,2] %>% # 443X78 * 78X1 
  as.data.frame() %>%
  dplyr::rename("bo_score_abs"="V1") %>%
  rownames_to_column(var="beadPosition") %>%
  inner_join(s2_clinical,by="beadPosition")

# score = score_abs %>% dplyr::rename("bo_score"= "bo_score_abs")

# correlation between ca/co and birth order score
ggplot(score, aes(x=as.factor(CaCo), y=bo_score, fill=as.factor(CaCo))) +
  geom_boxplot(outlier.size=0.2)+
  theme_classic() +
  xlab("self reported birth order") +
  ylab("birth order score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name="disease status",labels=c("0","1"))

t.test(score[score$CaCo==0,"bo_score"],
       score[score$CaCo==1,"bo_score"])

# correlation between birth order and birth order score
score %>%
  filter(!is.na(birth_order)) %>%
  ggplot(aes(x=as.factor(birth_order), y=bo_score, fill=as.factor(CaCo))) +
    geom_boxplot(outlier.size=0.2)+
    theme_classic() +
    xlab("case control status") +
    ylab("birth order score") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_discrete(name="disease status",labels=c("0","1"))
  
## scatter plot by group
score %>%
  mutate(CaCo=as.factor(CaCo)) %>%
  filter(!is.na(birth_order)) %>%
  ggplot(aes(x=birth_order, 
             y=bo_score,
             color=CaCo)) +
  geom_smooth(aes(group = CaCo),method='lm',se=FALSE) +
  geom_jitter() +
  theme_classic() +
  xlab("birth order") +
  ylab("birth order score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name="disease status",labels=c("0","1"))

## box plot by birth order
score %>%
  filter(!is.na(birth_order)) %>%
  ggplot(aes(x=as.factor(birth_order), y=bo_score)) +
    geom_boxplot(outlier.size=0.2)+
    theme_classic() +
    xlab("self reported birth order") +
    ylab("birth order score") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_discrete(name="disease status",labels=c("0","1"))


```

# sensitivity analysis

```{r}

setwd("/ccls/home/sli/ewas/birthorder/meta")

library(tidyverse)
library(data.table)

# 2A: including miscarriage/abortion

platform = "all"
race="all"
modelN="1A"
model <- fread(paste0(modelN,".",platform,".",race,".anno.txt")) %>%
  as.data.frame() %>%
  mutate(logP = -log10(P)) %>%
  dplyr::rename("Effect_1A"="Effect",
                "logP_1A"="logP") %>%
  dplyr::select(cpg,Effect_1A,logP_1A)

model2N="2A"
model2 <- fread(paste0(model2N,".",platform,".",race,".anno.txt")) %>%
  as.data.frame() %>%
  mutate(logP = -log10(P)) %>%
  dplyr::rename("Effect_2A"="Effect",
                "logP_2A"="logP")%>%
  dplyr::select(cpg,Effect_2A,logP_2A)

modelCompare = model %>%
  inner_join(model2,by="cpg")

ggplot(modelCompare,aes(x = Effect_1A, y = Effect_2A))+
  geom_point(alpha = 0.4,color='salmon',shape= 4) +
  geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
  xlab("Effect sizes,\n original model") + 
  ylab("Effect sizes,\n including miscarriags/abortions")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24))+
  theme_bw()
ggsave(paste0(model2N,".effect.size.sensitivity.png"),width = 6,height = 4)

ggplot(modelCompare,aes(x = logP_1A, y = logP_2A))+
  geom_point(alpha = 0.4,color='salmon',shape= 4) +
  geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
  xlab("log10 (P-value),\n original model") + 
  ylab("log10 (P-value),\n including miscarriags/abortions")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24))+
  theme_bw()
ggsave(paste0(model2N,".logP,sensitivity.png"),width = 6,height = 4)


# S3A: including + pre-pregnancy weight
model2N="S3A"
model2 <- fread(paste0(model2N,".",platform,".",race,".anno.txt")) %>%
  as.data.frame() %>%
  mutate(logP = -log10(P)) %>%
  dplyr::rename("Effect_2A"="Effect",
                "logP_2A"="logP")%>%
  dplyr::select(cpg,Effect_2A,logP_2A)

modelCompare = model %>%
  inner_join(model2,by="cpg")

ggplot(modelCompare,aes(x = Effect_1A, y = Effect_2A))+
  geom_point(alpha = 0.4,color='salmon',shape= 4) +
  geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
  xlab("Effect sizes,\n original model") + 
  ylab("Effect sizes,\n including maternal weight gain (g)")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24))+
  theme_bw()
ggsave(paste0(model2N,".effect.size.sensitivity.png"),width = 6,height = 4)

ggplot(modelCompare,aes(x = logP_1A, y = logP_2A))+
  geom_point(alpha = 0.4,color='salmon',shape= 4) +
  geom_abline(aes(intercept=0,slope=1),alpha=0.4,color="red")+
  xlab("log10 (P-value),\n original model") + 
  ylab("log10 (P-value),\n including maternal weight gain (g)")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24))+
  theme_bw()
ggsave(paste0(model2N,".logP.sensitivity.png"),width = 6,height = 4)

```

```{r what does abortion change?}

setwd("/ccls/home/sli/ewas/birthorder/meta")

library(tidyverse)
library(data.table)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
data("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
annoEPIC <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19 %>%
  getAnnotation %>%
  as.data.frame %>%
  rownames_to_column(var="cpg") %>%
  dplyr::select(cpg,chr,pos,Relation_to_Island,UCSC_RefGene_Name,
                Probe_maf,UCSC_RefGene_Group,CpG_maf)

platform = "all"
race="all"
modelN="1A"
model <- fread(paste0(modelN,".",platform,".",race,".anno.txt")) %>%
  as.data.frame() %>%
  mutate(logP = -log10(P)) %>%
  dplyr::rename("Effect_1A"="Effect",
                "logP_1A"="logP") %>%
  dplyr::select(cpg,Effect_1A,logP_1A)

model2N="2A"
model2 <- fread(paste0(model2N,".",platform,".",race,".anno.txt")) %>%
  as.data.frame() %>%
  mutate(logP = -log10(P)) %>%
  dplyr::rename("Effect_2A"="Effect",
                "logP_2A"="logP")%>%
  dplyr::select(cpg,Effect_2A,logP_2A)

modelCompare = model %>%
  inner_join(model2,by="cpg") %>%
  mutate(diffDis = abs(Effect_1A-Effect_2A)) %>%
  inner_join(annoEPIC,by="cpg") %>%
  arrange(desc(diffDis))%>%
  mutate(gene=gsub(";.*$","",UCSC_RefGene_Name)) %>%
  filter(gene!="")

modelCompareEnrich = modelCompare[1:100,] 
write.csv(modelCompareEnrich,"model.1a.2a.comp.enrich.csv")

# colon cancer? is this interesting?
```

